### 2023-12-22 docs/manuscript/manuscript.qmd
Primer commit

---
title: ""
bibliography: "misc/ref.bib"
csl: "misc/vancouver.csl"
format: 
  docx: 
    reference-doc: "ref.docx"
---

**Título**: [título]

**Title**: [title]

**Autores**: Autor,^1^ Autor,^2^ Autor,^3^ Autor,^4^ Autor,^5^ Autor.^6^ \*

::: {custom-style="Afiliation"}
^1^ Affiliation, location.

^2^ Affiliation, location.

^3^ Affiliation, location.

^4^ Affiliation, location.

^5^ Affiliation, location.

^6^ Affiliation, location.
:::

### \*Autor de correspondencia

Autor, Affiliation, location. e-mail: autor\@mail.com. 
Address: XXXX. Phone: XXXX.

\newpage

## Abstract

**Objective**: [...]. **Material and methods**: [...]. **Results**: [...]. **Conclusion**: [...].

**Keywords**: [...].

\newpage

# Introduction

<!---- Antecedentes históricos ---->
[...].

<!---- Conceptualización del problema ---->
[...].

<!---- Hipótesis y objetivos ---->
[...].

# Material and methods

## Study design

[...].

## Participants

[...].

## Procedures

[...].

## Assessments

### [Assessment 1]

[...].

### [Assessment 2]

[...].

### [Assessment 3]

[...].

## Statistical analysis

[...].

[...], using the *R* programming language for statistical computing, [@rlanguage].

# Results

[...].

[...].

# Discussion

[...].

[...].

[...].

# Conclusion

[...].

# Author Contributions

All authors listed have made a substantial, direct and intellectual contribution to the work and approved it for publication.

# Funding

[...].

# Institutional Review Board Statement

[...].

# Informed Consent Statement

[...]

# Data availability statement

The raw data supporting the conclusions of this article will be made available by the authors without undue reservation.

# Conflicts of interests

The authors declare that the research was conducted without any commercial or financial relationships construed as a potential conflict of interest.

# Referencias

<div id="refs"></div>

### 2023-12-22 data-raw/data-raw-process.R

@@ -1,0 +1,18 @@
+
+# Prepare workspace -------------------------------------------------------
+
+library(data.table)
+
+# Import dataset ----------------------------------------------------------
+
+raw <- fread(input = "data-raw/data-raw.csv")
+
+# Data wrangling ----------------------------------------------------------
+
+## Observar los valores unicos de cada variable
+lapply(raw, unique)
+
+
+
+# Data export -------------------------------------------------------------
+

### 2023-12-25 data-raw/data-raw-process.R
estandarización variables
- nombres estandarizados en col-names.txt
- rename en proceso

@@ -8,9 +8,25 @@ library(data.table)
 raw <- fread(input = "data-raw/data-raw.csv")
 
 # Data wrangling ----------------------------------------------------------
+colnames(raw)
+View(raw)
+ ##eliminar columnas repetidas sin valores
+raw_rename <- raw [, -"Complete?"]
+raw_rename <- raw_rename [, -"Survey Timestamp"]
+
+##renombrar columnas repetidas con valores
+names(raw_rename)[86] <- "bai_total"
+colnames(raw_rename)
+
+##renombrar columnas
+sapply(raw_rename, class)
+View(raw)
+
+
 
 ## Observar los valores unicos de cada variable
 lapply(raw, unique)
+
 
 
 

### 2024-02-08 docs/manuscript/manuscript.qmd
Introducción + Materiales y Métodos v1

@@
 
-**Título**: [título]
+**Título**: [La Aptitud Física se Relaciona con la Grasa Corporal y la Respuesta Autonómica Cardíaca al Ejercicio Físico en Personas Mayores]
 
-**Title**: [title]
+**Title**: [Physical Fitness is Related to Body Fat and Cardiac Autonomic Response to Exercise in Older People]
 
@@
 
+The process of aging is accompanied by a multitude of physiological changes that have the potential to significantly impact people's overall health and well-being.
+Body composition and cardiovascular function are particularly relevant, as they are closely associated with the aging process.
+In recent times, there has been a growing interest in comprehending the interplay between physical fitness, body composition, and cardiac autonomic response to exercise in older individuals, as these factors play pivotal roles in determining health outcomes within this age group.
+
 <!---- Antecedentes históricos ---->
-[...].
+
+Physical fitness is defined as the ability to perform daily tasks with vigor and without undue fatigue.
+It is a crucial component of healthy aging.
+The term encompasses a multitude of aspects, including cardiovascular endurance, muscular strength, flexibility, and balance.
+Numerous studies have demonstrated the positive impact of physical fitness on overall health and longevity, highlighting its role in reducing the risk of chronic diseases, improving quality of life, and improving functional independence in older adults.
+
+Moreover, the alteration of body composition, specifically the proportion of body fat, is accompanied by significant changes with age, resulting in a tendency towards an increase in adiposity and a decrease in lean muscle mass.
+Excessive body fat accumulation, particularly visceral adiposity, has been associated with a higher risk of cardiovascular disease, metabolic disorders, and functional limitations in older individuals.
+
+Autonomic nervous system (ANS) plays a critical role in regulating cardiovascular function, with sympathetic and parasympathetic branches exerting opposing effects on heart rate and vascular tone.
+Furthermore, during stressful situations, such as physical exercise, the ANS ensures a sufficient cardiac response to higher metabolic demands.
+The measurement of heart rate variability (HRV), serves as a non-invasive indicator of ANS activity and cardiovascular health.
+A reduction in HRV has been linked to various adverse outcomes, including cardiovascular events, mortality, and impaired exercise tolerance.
 
 <!---- Conceptualización del problema ---->
-[...].
+
+However, despite extensive research into physical fitness, body composition, and cardiovascular health in older adults, gaps persist in  understanding of their interrelationships and underlying mechanisms.
+One area of interest is the relationship between physical fitness, body fat percentage, and cardiac autonomic response to exercise in older individuals.
+Although previous studies have explored these factors independently, few have examined their collective influence on cardiovascular function during exercise in the aging population.
+
+The conceptualization of this study is based on the recognition of the interplay between physical fitness, body composition, and autonomic cardiovascular regulation in aging.
+We aim to answer the following research question: How does physical fitness relate to body fat percentage and cardiac autonomic response to exercise in older people?
+By elucidating these correlations, our objective is to gain valuable insights into the physiological mechanisms that underlie age-related modifications in cardiovascular function and identify potential interventions to enhance health outcomes in older adults.
 
 <!---- Hipótesis y objetivos ---->
-[...].
+
+We hypothesize that physical fitness will be inversely related to body fat percentage and positively associated with cardiac autonomic response to exercise in older individuals, based on existing literature and theoretical frameworks.
+In our opinion, it will be essential for this population to have an optimal cardiac autonomic response, characterized by balanced sympathetic and parasympathetic modulation.
+
+The primary objective of this study is to investigate the relationship between physical fitness, body fat percentage, and cardiac autonomic response to exercise in older people.
+We aimed to assess how variations in physical fitness and body composition influence autonomic cardiovascular regulation during low to moderate-intensity exercise, with implications for cardiovascular health and functional capacity in aging.
 
@@
 
-[...].
+A prospective, cross-sectional and correlational study was conducted on older people in one data collection session, to determine the associations between physical fitness, body composition and cardiac autonomic response to moderate-intensity exercise.
 
@@
 
-[...].
+XX older adults community-dwelling older adults were recruited and selected by non-probabilistic sampling.
+They were included if: 60 years age or older; residents of the Region of Magallanes and Chilean Antartica, Chile; and being able to understand the study aims and instructions.
+But they were excluded in case of: diagnosis of congenital heart disease; consumption of beta-blocker drugs, taking stimulant substances within 24 h before assessments session; motor or cognitive disability; or presence of pain during cardiac or physical assessments.
+
+All participating subjects gave their permission and provided informed consent before participation.
+The Ethics Committee of the University of Magallanes (N°XXX/YY/YEAR) approved this study following the regulations established by the Declaration of Helsinki on ethical principles in human beings.
 
@@
 
-[...].
+During recruitment,the participants were instructed to avoid the use of psychoactive substances for the 24 hours prior to the assessment, and to sleep for at least 7 hours during the previous night.
+Upon arrival, participants were informed about the study aims and risks associated with their participation, and the socio-demographic and medical information was collected during the initial interview.
+Then, Body composition parameters were measured using bioimpedance analysis.
 
-## Assessments
+Cardiac autonomic response to physical exercise was measured as previously validated for this age group (Castillo-Aguilar et al., 2023). The protocol is briefly described below:
 
-### [Assessment 1]
+HRV was recorded through a non-invasive chest band, immediately before, during, and immediatly after the execution of the two-minutes step test (TMST).
+Vital signs, including blood pressure, were monitored throughout the test, and participants' well-being was visually checked to ensure they were comfortable and mentally prepared.
+For resting HRV measurements (prior and posterior to TMST), the volunteers remained seated in a chair, while proper support of the feet and back, and avoiding talking during the recording was ensured.
+RR intervals were recorded continuously during the last 10 min of rest, and were subsequently analyzed for 5 min in each occasion.
+The breathing rate was spontaneous.
+As part of the protocol, it was ensured that the participant had a blood pressure less than 140/90 mmHg to start the HRV measurements.
 
-[...].
+Finally, after 15 minutes of resting from the TMST, the Short Physical Performance Battery (SPPB) was applied to the participants.
+Blood pressure, heart rate and general aspect of the subjects was verified before ending the session.
 
-### [Assessment 2]
+All the assessments were made between 09:00 and 11:00 hours of the morning.
+Measurements and tests were carried out by professional physiotherapists, ensuring the privacy and comfort of the subjects.
+The room temperature was set at 20°C and a white artificial lighting was used.
 
-[...].
+## Instruments
 
-### [Assessment 3]
+### [Short Physical Performance Battery]
 
-[...].
+SPPB consists of a physical test used to measure three components of physical fitness, described below (Guralnik et al., 1994):
+
+Balance: to achieve the maximal score (4 points), the subject should be able to stay balanced for at least 10 seconds in side-by-side stand (feet together), semi-tandem stand and tandem stand.
+
+Usual gait speed: the subjects are asked to walk normally a 4 meters distance.
+Two chances are given and the best of both is registered.
+To achieve the maximal score (4 points),they should obtain 4.82 sec.
+or less.
+
+Lower body fitness: The subjects are asked to perform 5 chair stands without using their arms and time upon complete is registered. A time lower or equal 11.19 sec. gives the maximal score (4 points).
+
+To obtain the final score, the summing of the three components is calculated, with a maximal total score of 12 points.
+
+### [Body Composition]
+
+Body mass (kg) and total body fat (%) were assessed by bioimpedance using the Tanita BC-558 Ironman Segmental Body Composition Monitor (Tanita Ironman, Arlington Heights, IL 60005, USA), with a concordance of 89.3% compared to the Dual X-ray Absorption test using standard measurement protocols (Mialich et al., 2011).
+
+### [Cardiac Autonomic Activity]
+
+Cardiac autonomic activity was assessed using RR interval recordings obtained through the Polar Team2 system (Polar®) application.
+Artifacts and ectopic heartbeats, which did not exceed 3% of the recorded data, were excluded [32].
+Time-domain parameters considered for analysis included the square root of the mean squared differences of successive RR intervals (RMSSD, expressed in ms) as an index of parasympathetic activity [33], and the standard deviation of RR intervals (SDNN), reflecting total variability encompassing both sympathetic and parasympathetic contributions to cardiac autonomic function [34,35].
+Additionally, the Stress Index (SI) and Parasympathetic and Sympathetic Nervous System Index (PNS and SNS, respectively) were calculated.
+The PNS Index, indicative of total vagal stimulation, was derived from mean R-R intervals, RMSSD, and Poincaré Plot Index SD1 in normalized units (linked to RMSSD), representing deviations from normal population averages [34,39].
+The SNS Index, reflecting total sympathetic stimulation, was derived from mean R-R intervals, Baevsky’s Stress Index (positively related to cardiovascular system stress and cardiac sympathetic activity), and the Poincaré Plot Index SD2 in normalized units (related to SDNN) with interpretation similar to the PNS Index [34,39].
+The SI serves as an indicator representing the ANS control system's workload [40], normalized by the square root of Baevsky’s SI [41].
+All analyses were conducted using Kubios HRV® software (Kuopio, Finland).
+### [Two-Minutes Step Test]
+
+The TMST is a sub-test from the Senior Fitness Test, demanding a low to moderate intensity (Rikli, 1999, Castillo-Aguilar et al., 2023).
+The test consists of a two-minute assessment designed to evaluate cardiorrespiratory fitness.
+Participants are instructed to raise their right knee to a marked point on a wall as many times as possible within the given time frame, ensuring that each raise reaches at least a 70° angle at the thigh-femoral joint.
 

### 2024-02-08 docs/manuscript/manuscript.qmd
Sección de resultados

@@
 
-**Título**: [La Aptitud Física se Relaciona con la Grasa Corporal y la Respuesta Autonómica Cardíaca al Ejercicio Físico en Personas Mayores]
+**Título**: La Aptitud Física se Relaciona con la Grasa Corporal y la Respuesta Autonómica Cardíaca al Ejercicio Físico en Personas Mayores
 
-**Title**: [Physical Fitness is Related to Body Fat and Cardiac Autonomic Response to Exercise in Older People]
+**Title**: Physical Fitness is Related to Body Fat and Cardiac Autonomic Response to Exercise in Older People
 
@@
 
-### [Short Physical Performance Battery]
+### Short Physical Performance Battery
 
@@
 
-### [Body Composition]
+### Body Composition
 
@@
 
-### [Cardiac Autonomic Activity]
+### Cardiac Autonomic Activity
 
@@
 All analyses were conducted using Kubios HRV® software (Kuopio, Finland).
-### [Two-Minutes Step Test]
+
+### Two-Minutes Step Test
 
@@
 
-[...].
-
 [...], using the *R* programming language for statistical computing, [@rlanguage].
@@
 
-[...].
+## Sample characteristics
 
-[...].
+Sociodemographic characteristic of the sample can be seen in table 1.
+
+## Body composition and autonomic response
+
+We observed that body fat percentage was linked to a lower SNS (r = -0.252, p = 0.030) and Stress index (r = 0.258, p = 0.027) response during the intervention.
+However, body fat percentage was positively correlated with time domain HRV during exercise, more specifically RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020).
+
+<!---
+#> cc_grasa_total_porcentaje, hrv_rmssd_peri, r = 0.253, p = 0.029
+#> cc_grasa_total_porcentaje, hrv_rmssd_peri, rho = 0.249, p = 0.032
+#> 
+#> cc_grasa_total_porcentaje, hrv_stress_peri, r = -0.258, p = 0.027
+#> cc_grasa_total_porcentaje, hrv_stress_peri, rho = -0.282, p = 0.015
+#> 
+#> cc_grasa_total_porcentaje, hrv_sns_peri, r = -0.252, p = 0.03
+#> cc_grasa_total_porcentaje, hrv_sns_peri, rho = -0.257, p = 0.027
+#> 
+#> cc_grasa_total_porcentaje, hrv_sdnn_peri, r = 0.269, p = 0.02
+#> cc_grasa_total_porcentaje, hrv_sdnn_peri, rho = 0.284, p = 0.014
+--->
+
+<!--- Figura con cuatro mostrando las correlaciones --->
+
+## Body composition and physical fitness
+
+In addition, a greater body fat percentage was negatively correlated with physical fitness (r = -0.273, p = 0.015).
+
+<!---
+#> cc_grasa_total_porcentaje, sppb_total, r = -0.273, p = 0.015
+#> cc_grasa_total_porcentaje, sppb_total, rho = -0.25, p = 0.026
+--->
+
+## Physical fitness and cardiac autonomic response to exercise
+
+Physical fitness was positively correlated with a greater simpathetic activity during exercise, which was reflected by SNS index (r = 0.313, p = 0.006), mean HR (r = 0.355, p = 0.002) and its inverse counterpart mean R-R (r = -0.35, p = 0.002).
+
+<!--- 
+#> hrv_sns_peri, sppb_total, r = 0.313, p = 0.006
+#> hrv_sns_peri, sppb_total, rho = 0.332, p = 0.004
+
+#> hrv_mean_hr_peri, sppb_total, r = 0.355, p = 0.002
+#> hrv_mean_hr_peri, sppb_total, rho = 0.375, p = 0.001
+
+#> hrv_mean_rr_peri, sppb_total, r = -0.35, p = 0.002
+#> hrv_mean_rr_peri, sppb_total, rho = -0.38, p = 0.001
+
+--->
+
+However, this appears to be accompanied by an proportional decrease in parasympathetic drive during exercise with increasing levels of physical fitness, which was denoted by the inverse correlation of the later with PNS index during (r = -0.375, p = 0.001) and after exercise (r = -0.229, p = 0.049), in addition to RMSSD during exercise (r = -0.294, p = 0.010)
+
+<!--- 
+#> hrv_pns_peri, sppb_total, r = -0.375, p = 0.001
+#> hrv_pns_peri, sppb_total, rho = -0.372, p = 0.001
+
+#> hrv_pns_post, sppb_total, r = -0.229, p = 0.049
+#> hrv_pns_post, sppb_total, rho = -0.238, p = 0.04
+
+#> hrv_rmssd_peri, sppb_total, r = -0.294, p = 0.01
+#> hrv_rmssd_peri, sppb_total, rho = -0.282, p = 0.014
+--->
+
+<!--- Figura 2 --->
 

### 2024-02-08 R/correlations.R

@@ -53,7 +53,8 @@ all_tag_pairs <- function() {
   apply(out, 1, any)
 }
 
-# Correlaciones -----------------------------------------------------------
+
+# Pearson r ---------------------------------------------------------------
 
 ## Primero realizaremos correlaciones entre los sujetos evaluados basalmente
 ## para explorar las relaciones latentes entre variables
@@ -73,9 +74,33 @@ cor_1 <- correlation(
 cor_1 <- as.data.table(cor_1)
 
 results_cor_1 <- cor_1[all_tag_pairs() & p < 0.05
-                       ][, list(Parameter1, Parameter2, r, p)]
+                       ][, list(Parameter1, Parameter2, r, p, n_Obs)]
 
 results_cor_1[order(Parameter1), list(
   paste0(Parameter1, ", ", Parameter2, ", r = ", round(r, 3), ", p = ", round(p, 3))
-)][, paste0(V1, collapse = "\n")] |> cat(file = "R/correlations_results.txt")
+)][, paste0(V1, collapse = "\n")] |> cat()
 
+
+# Spearman rho ------------------------------------------------------------
+
+cor_2 <- correlation(
+  data = mindfulness[
+    id_event_name == "Basal", 
+    j = .SD, 
+    .SDcols = grepl(
+      pattern = "^cc|^cv|^hrv|total$", 
+      x = names(mindfulness)
+    )
+  ],
+  p_adjust = "none", 
+  method = "spearman"
+)
+
+cor_2 <- as.data.table(cor_2)
+
+results_cor_2 <- cor_2[all_tag_pairs() & p < 0.05
+][, list(Parameter1, Parameter2, rho, p, n_Obs)]
+
+results_cor_2[order(Parameter1), list(
+  paste0(Parameter1, ", ", Parameter2, ", rho = ", round(rho, 3), ", p = ", round(p, 3))
+)][, paste0(V1, collapse = "\n")] |> cat()

### 2024-02-08 docs/manuscript/manuscript.qmd
Referencias añadidas introducción y mayteriales y metodos

--- 2024-02-08 docs/manuscript/manuscript.qmd
@@
 
-The process of aging is accompanied by a multitude of physiological changes that have the potential to significantly impact people's overall health and well-being.
-Body composition and cardiovascular function are particularly relevant, as they are closely associated with the aging process.
+The process of aging is accompanied by a multitude of physiological changes that have the potential to significantly impact people's overall health and well-being[@pmid:30185958], [@pmid:31154942].
+Body composition and cardiovascular function are particularly relevant, as they are closely associated with the aging process [@pmid:36948782], [@pmid:36443946], [@pmid:30065258], [@pmid:36918543], [@pmid:32483625].
 In recent times, there has been a growing interest in comprehending the interplay between physical fitness, body composition, and cardiac autonomic response to exercise in older individuals, as these factors play pivotal roles in determining health outcomes within this age group.
@@
 
-Cardiac autonomic response to physical exercise was measured as previously validated for this age group (Castillo-Aguilar et al., 2023). The protocol is briefly described below:
+Cardiac autonomic response to physical exercise was measured as previously validated for this age group [@ijerph20054456]. The protocol is briefly described below:
 
-HRV was recorded through a non-invasive chest band, immediately before, during, and immediatly after the execution of the two-minutes step test (TMST).
+HRV was recorded through a non-invasive chest band, immediately before, during, and immediately after the execution of the two-minutes step test (TMST).
 Vital signs, including blood pressure, were monitored throughout the test, and participants' well-being was visually checked to ensure they were comfortable and mentally prepared.
@@
 
-SPPB consists of a physical test used to measure three components of physical fitness, described below (Guralnik et al., 1994):
+SPPB consists of a physical test used to measure three components of physical fitness, described below [@Guralnik1994]:
 
@@
 
-Body mass (kg) and total body fat (%) were assessed by bioimpedance using the Tanita BC-558 Ironman Segmental Body Composition Monitor (Tanita Ironman, Arlington Heights, IL 60005, USA), with a concordance of 89.3% compared to the Dual X-ray Absorption test using standard measurement protocols (Mialich et al., 2011).
+Body mass (kg) and total body fat (%) were assessed by bioimpedance using the Tanita BC-558 Ironman Segmental Body Composition Monitor (Tanita Ironman, Arlington Heights, IL 60005, USA), with a concordance of 89.3% compared to the Dual X-ray Absorption test using standard measurement protocols [@Mialich2011].
 
@@
 
-The TMST is a sub-test from the Senior Fitness Test, demanding a low to moderate intensity (Rikli, 1999, Castillo-Aguilar et al., 2023).
+The TMST is a sub-test from the Senior Fitness Test, demanding a low to moderate intensity [@Rikli1999], [@ijerph20054456].
 The test consists of a two-minute assessment designed to evaluate cardiorrespiratory fitness.

### 2024-02-08 docs/manuscript/manuscript.qmd
arreglando algunas citas

@@
 The process of aging is accompanied by a multitude of physiological changes that have the potential to significantly impact people's overall health and well-being[@pmid:30185958], [@pmid:31154942].
-Body composition and cardiovascular function are particularly relevant, as they are closely associated with the aging process [@pmid:36948782], [@pmid:36443946], [@pmid:30065258], [@pmid:36918543], [@pmid:32483625].
+Body composition and cardiovascular function are particularly relevant, as they are closely associated with the aging process [@pmid:36948782; @pmid:36443946; @pmid:30065258; @pmid:36918543; @pmid:32483625].
 In recent times, there has been a growing interest in comprehending the interplay between physical fitness, body composition, and cardiac autonomic response to exercise in older individuals, as these factors play pivotal roles in determining health outcomes within this age group.

### 2024-02-18 docs/manuscript/manuscript.qmd
Secciones listas: - Abstract - Introduction - Study Design - Participants - Procedures - Measures - Discusion - Conclusion - References

@@
 ---
+**Title**: Physical Fitness is Related to Body Fat and Heart Rate Variability during Exercise in Community-Dwelling Older People.
 
-**Título**: La Aptitud Física se Relaciona con la Grasa Corporal y la Respuesta Autonómica Cardíaca al Ejercicio Físico en Personas Mayores
-
-**Title**: Physical Fitness is Related to Body Fat and Cardiac Autonomic Response to Exercise in Older People
-
-**Autores**: Autor,^1^ Autor,^2^ Autor,^3^ Autor,^4^ Autor,^5^ Autor.^6^ \*
+**Authors**: Autor,^1^ Autor,^2^ Autor,^3^ Autor,^4^ Autor,^5^ Autor.^6^ \*
 
@@
 
-### \*Autor de correspondencia
+### \*Corresponding author
 
@@
 
-**Objective**: [...]. **Material and methods**: [...]. **Results**: [...]. **Conclusion**: [...].
+**Objective**: to assess the correlation between physical fitness, body composition, and cardiac autonomic activity during moderate-intensity exercise in older people.
+**Material and methods**: an observational cross-sectional study was conducted in XX community-dwelling older people, recruited by non-probabilistic sampling.
+To obtain body composition parameters, a bioimpedance analysis was made, then, the Short Physical Performance Battery (SPPB) was used to assess physical fitness, and the Heart Rate Variability (HRV) was recorded before, during and after the Two-Minute Step Test (TMST), to assess cardiac autonomic response to physical exercise.
+**Results**: SPPB score was negatively related to body fat percentage (r = -0.273, p = 0.015), but positively related to SNS index during exercise (r = 0.313, p = 0.006).
+Also, a negative link existed between body fat and sympathetic activity: SNS index (r = -0.252, p = 0.030) and Stress index (r = 0.258, p = 0.027) during exercise.
+But a positive link with parasympathetic activity: RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020) **Conclusion**: the findings unveil a triangular relationship, where body fat percentage significantly impacts physical fitness and cardiac autonomic activity during exercise, while a greater sympathetic activity correlates with higher physical fitness.
+Additionally, physical fitness moderates the link between body fat and cardiac autonomic activity, emphasizing its importance for older adults' health.
+Further research is needed to explore psychological variables' moderating effects.
 
-**Keywords**: [...].
+**Keywords**: Aged, Physical Functional Performance, Exercise Test, Body Composition, Autonomic Nervous System
 
@@
 
-The process of aging is accompanied by a multitude of physiological changes that have the potential to significantly impact people's overall health and well-being[@pmid:30185958], [@pmid:31154942].
+The process of aging is accompanied by a multitude of physiological changes that have the potential to significantly impact people's overall health and well-being [@pmid:30185958; @pmid:31154942].
 Body composition and cardiovascular function are particularly relevant, as they are closely associated with the aging process [@pmid:36948782; @pmid:36443946; @pmid:30065258; @pmid:36918543; @pmid:32483625].
@@
 
-Physical fitness is defined as the ability to perform daily tasks with vigor and without undue fatigue.
-It is a crucial component of healthy aging.
-The term encompasses a multitude of aspects, including cardiovascular endurance, muscular strength, flexibility, and balance.
-Numerous studies have demonstrated the positive impact of physical fitness on overall health and longevity, highlighting its role in reducing the risk of chronic diseases, improving quality of life, and improving functional independence in older adults.
+Physical fitness has been classically defined as the ability to perform daily tasks with vigor and without undue fatigue, and with ample energy to enjoy leisure-time pursuits and to meet unforeseen emergencies [@pmid:3920716].
+Therefor, it is a crucial component of healthy aging.
+The term encompasses a multitude of aspects, typically including cardiorespiratory endurance, muscular endurance and strength, body composition, and flexibility [@pmid:3920711].
+Numerous studies have demonstrated the positive impact of physical fitness on overall health and longevity, highlighting its role in reducing the risk of chronic diseases, and improving quality of life and functional independence in older adults [@pmid:9177581; @pmid:29293447; @pmid:30768197; @pmid:36332759]
 
-Moreover, the alteration of body composition, specifically the proportion of body fat, is accompanied by significant changes with age, resulting in a tendency towards an increase in adiposity and a decrease in lean muscle mass.
-Excessive body fat accumulation, particularly visceral adiposity, has been associated with a higher risk of cardiovascular disease, metabolic disorders, and functional limitations in older individuals.
+Moreover, the alteration of body composition, specifically the proportion of body fat, is accompanied by significant changes with age, resulting in a tendency towards an increase in adiposity and a decrease in lean muscle mass [@pmid:35968789; @pmid:33673927].
+Excessive body fat accumulation, particularly visceral adiposity, has been associated with a higher risk of cardiovascular disease, metabolic disorders, and functional limitations in older individuals [@pmid:31301983; @pmid:33882682].
 
 Autonomic nervous system (ANS) plays a critical role in regulating cardiovascular function, with sympathetic and parasympathetic branches exerting opposing effects on heart rate and vascular tone.
-Furthermore, during stressful situations, such as physical exercise, the ANS ensures a sufficient cardiac response to higher metabolic demands.
-The measurement of heart rate variability (HRV), serves as a non-invasive indicator of ANS activity and cardiovascular health.
-A reduction in HRV has been linked to various adverse outcomes, including cardiovascular events, mortality, and impaired exercise tolerance.
+Furthermore, during stressful situations, such as physical exercise, the ANS ensures a sufficient cardiac response to higher metabolic demands [@pmid:16627049].
+The measurement of heart rate variability (HRV), serves as a non-invasive indicator of ANS activity and cardiovascular health [@pmid:37641564; @pmid:29033120].
+A reduction in rest HRV has been linked to various adverse outcomes, including cardiovascular events and morbidity, while insufficient HRV reduction during exercise may lead to impaired physical capacity [@pmid:29033120; @pmid:33390146; @pmid:36054204].
+
+However, HRV is not only influenced by physical and environmental stressors.
+Psychological factors such as anxiety and depression have also been linked to changes in ANS activity [@doi:10.1016/j.jad.2018.04.071; @pmid:35340102].
+Notably, older individuals may be more susceptible to these psychological factors, making it crucial to measure and control for their effects [@pmid:38007499; @pmid:25980510].
 
@@
 
-However, despite extensive research into physical fitness, body composition, and cardiovascular health in older adults, gaps persist in  understanding of their interrelationships and underlying mechanisms.
+Despite extensive research into physical fitness, body composition, and cardiovascular health in older adults, gaps persist in  understanding of their interrelationships and underlying mechanisms.
 One area of interest is the relationship between physical fitness, body fat percentage, and cardiac autonomic response to exercise in older individuals.
-Although previous studies have explored these factors independently, few have examined their collective influence on cardiovascular function during exercise in the aging population.
+Although previous studies have explored these factors independently, few have examined their collective influence on physical fitness in the aging population.
 
@@
 We aim to answer the following research question: How does physical fitness relate to body fat percentage and cardiac autonomic response to exercise in older people?
-By elucidating these correlations, our objective is to gain valuable insights into the physiological mechanisms that underlie age-related modifications in cardiovascular function and identify potential interventions to enhance health outcomes in older adults.
+By elucidating these correlations, we hope to gain valuable insights into the relationships that underlie age-related modifications in physical fitness and identify potential interventions to enhance health outcomes in older adults.
+Even more, we seek to understand the moderating effect of frequent psychological traits, such as geriatric depression and anxiety.
 
@@
 
-We hypothesize that physical fitness will be inversely related to body fat percentage and positively associated with cardiac autonomic response to exercise in older individuals, based on existing literature and theoretical frameworks.
+We hypothesize that physical fitness will be inversely related to body fat percentage and positively associated with cardiac autonomic response during exercise in older individuals, based on existing literature and theoretical frameworks.
 In our opinion, it will be essential for this population to have an optimal cardiac autonomic response, characterized by balanced sympathetic and parasympathetic modulation.
@@
 The primary objective of this study is to investigate the relationship between physical fitness, body fat percentage, and cardiac autonomic response to exercise in older people.
-We aimed to assess how variations in physical fitness and body composition influence autonomic cardiovascular regulation during low to moderate-intensity exercise, with implications for cardiovascular health and functional capacity in aging.
+Even more, We aimed to assess the moderating effect of psychological variables on these relationships, with implications for cardiovascular health and functional capacity in aging.
 
@@
 
-A prospective, cross-sectional and correlational study was conducted on older people in one data collection session, to determine the associations between physical fitness, body composition and cardiac autonomic response to moderate-intensity exercise.
+A prospective observational, cross-sectional study was conducted in older people in one data collection session, to determine the associations between physical fitness, body composition and cardiac autonomic response to moderate-intensity exercise.
 
@@
 
-XX older adults community-dwelling older adults were recruited and selected by non-probabilistic sampling.
-They were included if: 60 years age or older; residents of the Region of Magallanes and Chilean Antartica, Chile; and being able to understand the study aims and instructions.
-But they were excluded in case of: diagnosis of congenital heart disease; consumption of beta-blocker drugs, taking stimulant substances within 24 h before assessments session; motor or cognitive disability; or presence of pain during cardiac or physical assessments.
+XX community-dwelling older adults were recruited and selected by non-probabilistic sampling, from the Region of Magallanes and Chilean Antartica, Chile.
+They were included if: 65 years age or older; residents of the Region of Magallanes and Chilean Antartica, Chile; and being able to understand the study aims and instructions.
+But they were excluded in case of: diagnosis of congenital heart disease; consumption of beta-blocker drugs, taking stimulant substances within 24 h before assessments session; motor or cognitive disability; inability to understand instructions or written content; or presence of pain during cardiac or physical assessments.
 
@@
 
-During recruitment,the participants were instructed to avoid the use of psychoactive substances for the 24 hours prior to the assessment, and to sleep for at least 7 hours during the previous night.
+During recruitment, the participants were instructed to avoid the use of psychoactive substances for the 24 hours prior to the assessment, and to sleep for at least 7 hours during the previous night.
 Upon arrival, participants were informed about the study aims and risks associated with their participation, and the socio-demographic and medical information was collected during the initial interview.
-Then, Body composition parameters were measured using bioimpedance analysis.
+Then, Body composition parameters were measured using bioimpedance analysis, and psychological questionnaires for anxiety and depression screening were applied by a supervised psychology undergraduate student.
 
@@
 Two chances are given and the best of both is registered.
-To achieve the maximal score (4 points),they should obtain 4.82 sec.
+To achieve the maximal score (4 points), they should obtain 4.82 sec.
 or less.
@@
 
-Cardiac autonomic activity was assessed using RR interval recordings obtained through the Polar Team2 system (Polar®) application.
-Artifacts and ectopic heartbeats, which did not exceed 3% of the recorded data, were excluded [32].
-Time-domain parameters considered for analysis included the square root of the mean squared differences of successive RR intervals (RMSSD, expressed in ms) as an index of parasympathetic activity [33], and the standard deviation of RR intervals (SDNN), reflecting total variability encompassing both sympathetic and parasympathetic contributions to cardiac autonomic function [34,35].
+Cardiac autonomic activity was assessed using RR interval recordings (HRV) obtained through the Polar Team2 system (Polar®) application.
+Artifacts and ectopic heartbeats, which did not exceed 3% of the recorded data, were excluded [@malik1996heart].
+Time-domain parameters considered for analysis included the square root of the mean squared differences of successive RR intervals (RMSSD, expressed in ms) as an index of parasympathetic activity [@pmid:20033207], and the standard deviation of RR intervals (SDNN), reflecting total variability encompassing both sympathetic and parasympathetic contributions to cardiac autonomic function [@pmid:9401419; @buchheit2006cardiac].
 Additionally, the Stress Index (SI) and Parasympathetic and Sympathetic Nervous System Index (PNS and SNS, respectively) were calculated.
-The PNS Index, indicative of total vagal stimulation, was derived from mean R-R intervals, RMSSD, and Poincaré Plot Index SD1 in normalized units (linked to RMSSD), representing deviations from normal population averages [34,39].
-The SNS Index, reflecting total sympathetic stimulation, was derived from mean R-R intervals, Baevsky’s Stress Index (positively related to cardiovascular system stress and cardiac sympathetic activity), and the Poincaré Plot Index SD2 in normalized units (related to SDNN) with interpretation similar to the PNS Index [34,39].
-The SI serves as an indicator representing the ANS control system's workload [40], normalized by the square root of Baevsky’s SI [41].
+The PNS Index, indicative of total vagal stimulation, was derived from mean R-R intervals, RMSSD, and Poincaré Plot Index SD1 in normalized units (linked to RMSSD), representing deviations from normal population averages [@pmid:9401419; @GSID:XkAQwTGLgNsJ].
+The SNS Index, reflecting total sympathetic stimulation, was derived from mean R-R intervals, Baevsky’s Stress Index (positively related to cardiovascular system stress and cardiac sympathetic activity), and the Poincaré Plot Index SD2 in normalized units (related to SDNN) with interpretation similar to the PNS Index [@pmid:9401419; @GSID:XkAQwTGLgNsJ].
+The SI serves as an indicator representing the ANS control system's workload [@doi:10.1159/000513781], normalized by the square root of Baevsky’s SI [@2009MethodicalRU].
 All analyses were conducted using Kubios HRV® software (Kuopio, Finland).
@@
 
-The TMST is a sub-test from the Senior Fitness Test, demanding a low to moderate intensity [@Rikli1999], [@ijerph20054456].
+The TMST is a sub-test from the Senior Fitness Test, demanding a low to moderate intensity [@Rikli1999; @ijerph20054456].
 The test consists of a two-minute assessment designed to evaluate cardiorrespiratory fitness.
 Participants are instructed to raise their right knee to a marked point on a wall as many times as possible within the given time frame, ensuring that each raise reaches at least a 70° angle at the thigh-femoral joint.
+
+### Geriatric Depression Scale
+
+The 30-questions Geriatric Depression Scale (GDS-30) was employed to assess subject's depressivity [@pmid:27633329].
+It consists of a 30 dichotomous questionnaire, where participants are asked about their past-week feelings in reference to depressive symptoms (for instance, "Do you feel that your life is empty?"), with higher scores (i.
+e., "yes" responses) representing more depressive symptoms.
+It was first developed by Yesavage and colleagues [@pmid:7183759] and is currently widely used [@pmid:34922595].
+The spanish version, employed in this study, has been previously validated, with a Cronbach alpha coefficient of 0.82 [@pmid:11921157].
+
+### Beck Anxiety Inventory
+
+The Beck anxiety Inventory was employed to assess anxiety symptoms.
+It consists of a 21-items questionnaire based on usual anxiety symptoms and a 4-options Likert scale from 0 ("Not at all") to 3 ("Severely"), meaning the severity reported by the subject in each one.
+It was originally developed by Beck and colleagues [@pmid:3204199].
+In this study, the spanish version is used, which has demonstrated a high internal consistency in older people (α = 0.94) [@RODRIGUEZREYNALDO2001353].
 
@@
 We observed that body fat percentage was linked to a lower SNS (r = -0.252, p = 0.030) and Stress index (r = 0.258, p = 0.027) response during the intervention.
-However, body fat percentage was positively correlated with time domain HRV during exercise, more specifically RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020).
+In addition, body fat percentage was positively correlated with time domain HRV during exercise, more specifically RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020).
 
@@
 
-Physical fitness was positively correlated with a greater simpathetic activity during exercise, which was reflected by SNS index (r = 0.313, p = 0.006), mean HR (r = 0.355, p = 0.002) and its inverse counterpart mean R-R (r = -0.35, p = 0.002).
+Physical fitness was positively correlated with a greater sympathetic activity during exercise, which was reflected by SNS index (r = 0.313, p = 0.006), mean HR (r = 0.355, p = 0.002) and its inverse counterpart mean R-R (r = -0.35, p = 0.002).
 
@@
 
-[...].
+This study aimed to investigate the relationship between physical fitness, body fat percentage, and cardiac autonomic response to exercise in older people.
+Through the correlational analysis, we aimed to elucidate the  interplay between these factors and their implications for cardiac autonomic function during physical efforts.
 
-[...].
+It was hypothesized that a higher body fat percentage would be inversely related to  physical aptitude.
+Our results aligns with this hypothesis and previous research regarding body composition changes and physical performance decrease during aging [@pmid:35832415; @pmid:36256894].
 
-[...].
+Furthermore, the increase in body fat percentage was related to a reduced sympathetic activity response and a higher parasympathetic activity during TMST, as hypothesized (i.
+e., SNS index, SI, SDNN and RMSSD).
+Throughout the course of physical efforts, ANS is expected to shift into a sympathetic predominance, ensuring a sufficient response to higher metabolic demands [@pmid:33246163].
+However, aged individuals may present an impaired ANS response to physical exercise [@ijerph20054456].
+In that sense, our results support that body composition may play a role in that response, opening opportunities for targeted interventions.
+
+Nonetheless, the moderating effect of physical fitness on the relationship proposed above, suggests that more physically active individuals may counteract the negative effect of a high body fat percentage on cardiac autonomic response to exercise.
+This result highlights the importance of promoting a physically active lifestyle and favor the maintenance of physical capacity, regardless of nutritional status [@pmid:33402165; @pmid:30002462].
+
+Regarding physical fitness and cardiac autonomic response to exercise, as expected, our results evidenced that individuals with a higher sympathetic activity during exercise, as measured by SNS index, had a greater physical fitness while the opposite observation was made for parasympathetic activity (i.
+e., PNS index and RMSSD).
+Previous research had demonstrated the utility of HRV decrease as an index of cardiorespiratory fitness [@pmid:36054204], supporting the hypothesis that a higher sympathetic activity facilitates the performance during physical moderate effort.
+Even more, it has been previously demonstrated that physically active individuals may have an enhanced autonomic response to exercise, strengthening this relationship [@pmid:28721030; @pmid:24799513].
+
+Our results demonstrate the interplay between physical fitness, body fat, and cardiac autonomic response to exercise in older individuals.
+Physical fitness is negatively affected by higher body fat percentage , while also being associated with impaired cardiac autonomic response to exercise, characterized by decreased sympathetic activity and increased parasympathetic tone.
+Additionally, this altered autonomic response is related to lower physical performance in older adults, both variables being important predictors of health outcomes in this population [@pmid:30627571; @pmid:30104034].
+Overall, our findings underscore the triangular relationship among these variables.
+
+However, our results did not evidence a moderating effect of GDS and BAI scores, challenging initial hypothesis.
+This finding suggests a nuanced interplay between psychological variables and physical/physiological measures in older people, underscoring the need for further research to elucidate underlying mechanisms and variables beyond depression and anxiety.
+While the lack of moderating effects does not diminish the significance of psychological factors in aging, it highlights the complexity of their influence.
+
+Despite these findings, this study is not without limitations.
+In first place, our cross-sectional design prevents us from making objective inferences about the direction of the relationships.
+In second place, a relatively small sample was included, limiting generalizability and statistical power of the analysis.
+It should also be noted that the relationships between physical fitness and cardiac autonomic response to exercise were not made within the same instrument, due to the possible lack of SPPB to generate an adequate physical stress, and of TMST to inform general physical fitness of the participants.
+Further research could investigate the impact of cardiac autonomic response on performance during the same task, in this age group.
 
@@
 
-[...].
+The results highlighted a triangular relationship between the variables.
+Body fat percentage was importantly related to both physical fitness and cardiac autonomic activity during exercise, while sympathetic activity was associated with a grater physical fitness in older people.
+Furthermore, physical fitness may moderate the associations between body fat and cardiac autonomic activity, highlighting its importance in older people's health.
+However, further research is needed to elucidate moderating effects of psychological variables in this field.
 

### 2024-02-19 docs/manuscript/manuscript.qmd
modificaciones mínimas

@@
 
-We hypothesize that physical fitness will be inversely related to body fat percentage and positively associated with cardiac autonomic response during exercise in older individuals, based on existing literature and theoretical frameworks.
-In our opinion, it will be essential for this population to have an optimal cardiac autonomic response, characterized by balanced sympathetic and parasympathetic modulation.
+We hypothesize that physical fitness will be inversely related to body fat percentage and associated with cardiac autonomic response during exercise in older individuals.
+In our opinion, it will be essential for this population to have an optimal cardiac autonomic response, characterized by a shift towards sympathetic activity predominance.
 
@@
 
-However, our results did not evidence a moderating effect of GDS and BAI scores, challenging initial hypothesis.
+However, our results did not evidence a moderating effect of GDS-30 and BAI scores, challenging initial hypothesis.
 This finding suggests a nuanced interplay between psychological variables and physical/physiological measures in older people, underscoring the need for further research to elucidate underlying mechanisms and variables beyond depression and anxiety.

### 2024-02-21 docs/manuscript/manuscript.qmd
update on manuscript, tables and analysis
- descriptive tables
- mediation analysis on statistical analyses

--- 2024-02-08 docs/manuscript/manuscript.qmd
@@
 
-**Título**: La Aptitud Física se Relaciona con la Grasa Corporal y la Respuesta Autonómica Cardíaca al Ejercicio Físico en Personas Mayores
+**Title**: Physical fitness is related to body fat and cardiac autonomic response to exercise in older people.
 
-**Title**: Physical Fitness is Related to Body Fat and Cardiac Autonomic Response to Exercise in Older People
-
-**Autores**: Autor,^1^ Autor,^2^ Autor,^3^ Autor,^4^ Autor,^5^ Autor.^6^ \*
+**Authors**: Autor,^1^ Autor,^2^ Autor,^3^ Autor,^4^ Autor,^5^ Autor.^6^ \*
 
@@
 
-### \*Autor de correspondencia
+### \*Correspondence
 
@@
 
-<!---- Antecedentes históricos ---->
+<!---- Historical precedents ---->
 
@@
 
-<!---- Conceptualización del problema ---->
+<!---- Conceptualization of the issue ---->
 
@@
 
-<!---- Hipótesis y objetivos ---->
+<!---- Hypothesis and main objectives ---->
 
@@
 
-### Short Physical Performance Battery
+### Short physical performance battery
 
@@
 
-### Body Composition
+### Body composition
 
@@
 
-### Cardiac Autonomic Activity
+### Cardiac autonomic activity
 
@@
 
-### Two-Minutes Step Test
+### Two-minutes step test
 
-The TMST is a sub-test from the Senior Fitness Test, demanding a low to moderate intensity [@Rikli1999], [@ijerph20054456].
+The TMST is a sub-test from the Senior Fitness Test, demanding a low to moderate intensity [@Rikli1999; @ijerph20054456].
 The test consists of a two-minute assessment designed to evaluate cardiorrespiratory fitness.
@@
 
-[...], using the *R* programming language for statistical computing, [@rlanguage].
+We used mean and standard deviation (SD) to describe continuous variables, absolute and relative frequencies to describe discrete variables.
+In order to assess the realtionship between variables, we used Pearson's product-moment correlation (r), and to assess differences between groups we used standardized mean difference (SMD) and 95% confidence intervals (CI~95%~).
+
+As a way of controlling for the influence of psycho-physiological variables on the cardiac autonomic response of the subjects to exercise, we assess the average causal mediaton effect (ACME), the average direct effect (ADE) of the main effects after taking into account the effect of moderator variables into the observed relationships and the proportion of the effect that its mediated by these variables.
+To assess the significance of moderating variables, we used non parametric bootstrapping by using Monte Carlo draws, and bias-corrected and accelerated CI~95%~.
+
+We defined a type I error rate at 5% ($\alpha$ = 0.05) as our threshold for null hypothesis significance testing.
+All analyses were computed using the *R* programming language for statistical computing, [@rlanguage].
 
@@
 
-Sociodemographic characteristic of the sample can be seen in table 1.
+The sample assessed consisted on 81 adults (age: 71.1 ± 6.2 years old,  BMI: 31 ± 6.2 kg/m^2^), of whom 67 (82.7%) were females.
+Sociodemographic and anthropometric characteristics of the sample can be seen in table 1.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+library(flextable)
+readRDS(file = "tables/tbl-descriptives.RDS") |> 
+  font(part = "all", fontname = "Times")
+```
 

### 2024-02-21 data-raw/data-raw-process.R

@@ -119,11 +119,6 @@ raw[, tm_categoria := factor(tm_categoria,
                              levels = c("BAJO RANGO", "NORMAL", "ALTO RANGO"),
                              labels = c("Bajo", "Normal", "Alto"))]
 
-#> cc_imc_categoria <- standardize as ordered factor
-raw[, cc_imc_categoria := factor(cc_imc_categoria, 
-                                 levels = c("Peso normal", "Sobrepeso", "Obeso"),
-                                 ordered = TRUE)]
-
 ## Function to change missformated numbers into numeric format
 text_to_number <- function(x, remove_dash = TRUE) {
   x[grepl("[A-Za-z]", x) | nchar(x) == 0] <- NA
@@ -148,6 +143,21 @@ raw[, cc_masa_muscular_brazo_derecho := text_to_number(cc_masa_muscular_brazo_de
 
 #> cc_peso_corporal <- change to number
 raw[, cc_peso_corporal := text_to_number(cc_peso_corporal)]
+
+#> cc_talla <- Remove outlier
+raw[cc_talla == "", cc_talla := NA]
+raw[cc_talla %like% "\\,", cc_talla := gsub("\\,", "\\.", cc_talla)]
+raw[, cc_talla := as.numeric(cc_talla)]
+raw[cc_talla < 2, cc_talla := cc_talla * 100]
+
+#> cc_imc <- recalculate
+raw[c(68,187), cc_peso_corporal := round(cc_imc * (cc_talla / 100), 1)]
+raw[, cc_imc := round(cc_peso_corporal / ((cc_talla / 100) ^ 2), 2)]
+
+#> cc_imc_categoria <- Recompute labels
+raw[, cc_imc_categoria := cut(cc_imc, 
+                              breaks = c(0, 18.5, 24.9, 29.9, Inf), 
+                              labels = c("Infrapeso", "Normopeso", "Sobrepeso", "Obeso"))]
 
 #> cc_masa_osea_porcentaje <- change to number
 raw[, cc_masa_osea_porcentaje := text_to_number(cc_masa_osea_porcentaje)]
@@ -233,7 +243,7 @@ z_score <- function(i) {
 std_vars <- mindfulness[, lapply(.SD, z_score), .SDcols = grepl("^id|^cc|^hrv|^cv", names(mindfulness))]
 
 ## Select those observations with 3 SD above or below the mean
[…]

### 2024-02-22 docs/manuscript/manuscript.qmd
Cambios en título y adaptación de secciones: - ":a cross-sectional study" - "Setting", "Methods", "Aims and Study Design"

--- 2024-02-19 docs/manuscript/manuscript.qmd
@@
 ---
-**Title**: Physical Fitness is Related to Body Fat and Heart Rate Variability during Exercise in Community-Dwelling Older People.
+**Title**: Physical Fitness is Related to Body Fat and Heart Rate Variability during Exercise in Community-Dwelling Older People: a cross-sectional study.
 
@@
 
-**Objective**: to assess the correlation between physical fitness, body composition, and cardiac autonomic activity during moderate-intensity exercise in older people.
-**Material and methods**: an observational cross-sectional study was conducted in XX community-dwelling older people, recruited by non-probabilistic sampling.
-To obtain body composition parameters, a bioimpedance analysis was made, then, the Short Physical Performance Battery (SPPB) was used to assess physical fitness, and the Heart Rate Variability (HRV) was recorded before, during and after the Two-Minute Step Test (TMST), to assess cardiac autonomic response to physical exercise.
+**Background**: Physical fitness, body composition, and the autonomic response to exercise are important predictors of health-related outcomes in older adults, but their interplay is not fully understood, therefore, this study aims to assess the associations between these variables and possible mediating effects.
+**Methods**: an observational cross-sectional study was conducted in XX community-dwelling older people, recruited by non-probabilistic sampling.
+To obtain body composition parameters, a bio-impedance analysis was made, then, the Short Physical Performance Battery (SPPB) was used to assess physical fitness, and the Heart Rate Variability (HRV) was recorded before, during and after the Two-Minute Step Test (TMST), to assess cardiac autonomic response to physical exercise.
+Geriatric depression and anxiety symptoms were assessed as secondary measures.
+Correlation and mediating effects analyses were carried out.
 **Results**: SPPB score was negatively related to body fat percentage (r = -0.273, p = 0.015), but positively related to SNS index during exercise (r = 0.313, p = 0.006).
-Also, a negative link existed between body fat and sympathetic activity: SNS index (r = -0.252, p = 0.030) and Stress index (r = 0.258, p = 0.027) during exercise.
+Also, a negative link existed between body fat and sympathetic activity: SNS index (r = -0.252, p = 0.030) and Stress index (r = 0.258, p = 0.027) during TMST.
 But a positive link with parasympathetic activity: RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020) **Conclusion**: the findings unveil a triangular relationship, where body fat percentage significantly impacts physical fitness and cardiac autonomic activity during exercise, while a greater sympathetic activity correlates with higher physical fitness.
@@
 
-Despite extensive research into physical fitness, body composition, and cardiovascular health in older adults, gaps persist in  understanding of their interrelationships and underlying mechanisms.
+Despite extensive research into physical fitness, body composition, and cardiovascular health in older adults, gaps persist in the understanding of their interrelationships and underlying mechanisms.
 One area of interest is the relationship between physical fitness, body fat percentage, and cardiac autonomic response to exercise in older individuals.
-Although previous studies have explored these factors independently, few have examined their collective influence on physical fitness in the aging population.
+Although previous studies have explored these factors independently, this study explore their collective influence on physical fitness in the aging population.
 
@@
 We hypothesize that physical fitness will be inversely related to body fat percentage and associated with cardiac autonomic response during exercise in older individuals.
-In our opinion, it will be essential for this population to have an optimal cardiac autonomic response, characterized by a shift towards sympathetic activity predominance.
+In our opinion, it will be essential for this population to have an optimal cardiac autonomic response, characterized by a shift towards sympathetic activity predominance, in order to achieve a grater performance during physical fitness testing.
 
@@
 
-# Material and methods
+# Methods
 
-## Study design
+## Aims and Study design
 
-A prospective observational, cross-sectional study was conducted in older people in one data collection session, to determine the associations between physical fitness, body composition and cardiac autonomic response to moderate-intensity exercise.
+A prospective observational, cross-sectional study was conducted  in one data collection session, to determine the associations between physical fitness, body composition and cardiac autonomic response to moderate-intensity exercise.
+
+## Setting
+
+This study was conducted at the Centro Asistencial Docente e Investigación (Assistance, Teaching, and Research Center), belonging to the University of Magallanes (CADI-UMAG), in the city of Punta Arenas, Chile.
+All the assessments were made between 09:00 and 11:00 hours of the morning.
+The privacy and comfort of the subjects was ensured, the room temperature was set at 20°C and a white artificial lighting was used.
 
@@
 
-All the assessments were made between 09:00 and 11:00 hours of the morning.
-Measurements and tests were carried out by professional physiotherapists, ensuring the privacy and comfort of the subjects.
-The room temperature was set at 20°C and a white artificial lighting was used.
+Physical and Physiological assessments, as well as the initial interview, were made by professional physiotherapists.
 
@@
 
-We observed that body fat percentage was linked to a lower SNS (r = -0.252, p = 0.030) and Stress index (r = 0.258, p = 0.027) response during the intervention.
+We observed that body fat percentage was linked to a lower SNS (r = -0.252, p = 0.030) and Stress index (r = 0.258, p = 0.027) response during the TSMT.
 In addition, body fat percentage was positively correlated with time domain HRV during exercise, more specifically RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020).
@@
 
-Despite these findings, this study is not without limitations.
-In first place, our cross-sectional design prevents us from making objective inferences about the direction of the relationships.
+In addition to the potential strengths of this cross-sectional study, it is not without limitations.
+In first place, our study design prevents us from making objective inferences about the direction of the relationships, highlighting the relevance of longitudinal research exploring causal relations between our variables and underlying mechanisms.
 In second place, a relatively small sample was included, limiting generalizability and statistical power of the analysis.

### 2024-02-22 docs/manuscript/manuscript.qmd
Cambios menores: -N añadido al abstract y Participants -Limitación añadida a discusión

@@
 **Background**: Physical fitness, body composition, and the autonomic response to exercise are important predictors of health-related outcomes in older adults, but their interplay is not fully understood, therefore, this study aims to assess the associations between these variables and possible mediating effects.
-**Methods**: an observational cross-sectional study was conducted in XX community-dwelling older people, recruited by non-probabilistic sampling.
+**Methods**: an observational cross-sectional study was conducted in 81 community-dwelling older people, recruited by non-probabilistic sampling.
 To obtain body composition parameters, a bio-impedance analysis was made, then, the Short Physical Performance Battery (SPPB) was used to assess physical fitness, and the Heart Rate Variability (HRV) was recorded before, during and after the Two-Minute Step Test (TMST), to assess cardiac autonomic response to physical exercise.
@@
 
-XX community-dwelling older adults were recruited and selected by non-probabilistic sampling, from the Region of Magallanes and Chilean Antartica, Chile.
+81 community-dwelling older adults were recruited and selected by non-probabilistic sampling, from the Region of Magallanes and Chilean Antartica, Chile.
 They were included if: 65 years age or older; residents of the Region of Magallanes and Chilean Antartica, Chile; and being able to understand the study aims and instructions.
@@
 In second place, a relatively small sample was included, limiting generalizability and statistical power of the analysis.
+Furthermore, the sample consisted mainly of women (82.7%), limiting our ability to make sex comparisons and decreasing the representativeness of the sample.
 It should also be noted that the relationships between physical fitness and cardiac autonomic response to exercise were not made within the same instrument, due to the possible lack of SPPB to generate an adequate physical stress, and of TMST to inform general physical fitness of the participants.

### 2024-02-22 docs/manuscript/manuscript.qmd
update on statistical analyses

--- 2024-02-22 docs/manuscript/manuscript.qmd
@@
 ---
-**Title**: Physical Fitness is Related to Body Fat and Heart Rate Variability during Exercise in Community-Dwelling Older People: a cross-sectional study.
+**Title**: Physical Fitness is Related to Body Fat and Heart Rate Variability during Exercise in Community-Dwelling Older People.
 
@@
 
-**Background**: Physical fitness, body composition, and the autonomic response to exercise are important predictors of health-related outcomes in older adults, but their interplay is not fully understood, therefore, this study aims to assess the associations between these variables and possible mediating effects.
-**Methods**: an observational cross-sectional study was conducted in XX community-dwelling older people, recruited by non-probabilistic sampling.
-To obtain body composition parameters, a bio-impedance analysis was made, then, the Short Physical Performance Battery (SPPB) was used to assess physical fitness, and the Heart Rate Variability (HRV) was recorded before, during and after the Two-Minute Step Test (TMST), to assess cardiac autonomic response to physical exercise.
-Geriatric depression and anxiety symptoms were assessed as secondary measures.
-Correlation and mediating effects analyses were carried out.
+**Objective**: to assess the correlation between physical fitness, body composition, and cardiac autonomic activity during moderate-intensity exercise in older people.
+**Material and methods**: an observational cross-sectional study was conducted in XX community-dwelling older people, recruited by non-probabilistic sampling.
+To obtain body composition parameters, a bioimpedance analysis was made, then, the Short Physical Performance Battery (SPPB) was used to assess physical fitness, and the Heart Rate Variability (HRV) was recorded before, during and after the Two-Minute Step Test (TMST), to assess cardiac autonomic response to physical exercise.
 **Results**: SPPB score was negatively related to body fat percentage (r = -0.273, p = 0.015), but positively related to SNS index during exercise (r = 0.313, p = 0.006).
-Also, a negative link existed between body fat and sympathetic activity: SNS index (r = -0.252, p = 0.030) and Stress index (r = 0.258, p = 0.027) during TMST.
+Also, a negative link existed between body fat and sympathetic activity: SNS index (r = -0.252, p = 0.030) and Stress index (r = 0.258, p = 0.027) during exercise.
 But a positive link with parasympathetic activity: RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020) **Conclusion**: the findings unveil a triangular relationship, where body fat percentage significantly impacts physical fitness and cardiac autonomic activity during exercise, while a greater sympathetic activity correlates with higher physical fitness.
@@
 
-Despite extensive research into physical fitness, body composition, and cardiovascular health in older adults, gaps persist in the understanding of their interrelationships and underlying mechanisms.
+Despite extensive research into physical fitness, body composition, and cardiovascular health in older adults, gaps persist in  understanding of their interrelationships and underlying mechanisms.
 One area of interest is the relationship between physical fitness, body fat percentage, and cardiac autonomic response to exercise in older individuals.
-Although previous studies have explored these factors independently, this study explore their collective influence on physical fitness in the aging population.
+Although previous studies have explored these factors independently, few have examined their collective influence on physical fitness in the aging population.
 
@@
 We hypothesize that physical fitness will be inversely related to body fat percentage and associated with cardiac autonomic response during exercise in older individuals.
-In our opinion, it will be essential for this population to have an optimal cardiac autonomic response, characterized by a shift towards sympathetic activity predominance, in order to achieve a grater performance during physical fitness testing.
+In our opinion, it will be essential for this population to have an optimal cardiac autonomic response, characterized by a shift towards sympathetic activity predominance.
 
@@
 
-# Methods
+# Material and methods
 
-## Aims and Study design
+## Study design
 
-A prospective observational, cross-sectional study was conducted  in one data collection session, to determine the associations between physical fitness, body composition and cardiac autonomic response to moderate-intensity exercise.
-
-## Setting
-
-This study was conducted at the Centro Asistencial Docente e Investigación (Assistance, Teaching, and Research Center), belonging to the University of Magallanes (CADI-UMAG), in the city of Punta Arenas, Chile.
-All the assessments were made between 09:00 and 11:00 hours of the morning.
-The privacy and comfort of the subjects was ensured, the room temperature was set at 20°C and a white artificial lighting was used.
+A prospective observational, cross-sectional study was conducted in older people in one data collection session, to determine the associations between physical fitness, body composition and cardiac autonomic response to moderate-intensity exercise.
 
@@
 
-XX community-dwelling older adults were recruited and selected by non-probabilistic sampling, from the Region of Magallanes and Chilean Antartica, Chile.
-They were included if: 65 years age or older; residents of the Region of Magallanes and Chilean Antartica, Chile; and being able to understand the study aims and instructions.
+XX community-dwelling older adults were recruited and selected by non-probabilistic sampling, from the Region of Magallanes and Chilean Antarctica, Chile.
+They were included if: 65 years age or older; residents of the Region of Magallanes and Chilean Antarctica, Chile; and being able to understand the study aims and instructions.
 But they were excluded in case of: diagnosis of congenital heart disease; consumption of beta-blocker drugs, taking stimulant substances within 24 h before assessments session; motor or cognitive disability; inability to understand instructions or written content; or presence of pain during cardiac or physical assessments.
@@
 
-Physical and Physiological assessments, as well as the initial interview, were made by professional physiotherapists.
+All the assessments were made between 09:00 and 11:00 hours of the morning.
+Measurements and tests were carried out by professional physiotherapists, ensuring the privacy and comfort of the subjects.
+The room temperature was set at 20°C and a white artificial lighting was used.
 
@@
 The TMST is a sub-test from the Senior Fitness Test, demanding a low to moderate intensity [@Rikli1999; @ijerph20054456].
-The test consists of a two-minute assessment designed to evaluate cardiorrespiratory fitness.
+The test consists of a two-minute assessment designed to evaluate cardiorespiratory fitness.
 Participants are instructed to raise their right knee to a marked point on a wall as many times as possible within the given time frame, ensuring that each raise reaches at least a 70° angle at the thigh-femoral joint.
@@
 
-The 30-questions Geriatric Depression Scale (GDS-30) was employed to assess subject's depressivity [@pmid:27633329].
+The 30-questions Geriatric Depression Scale (GDS-30) was employed to assess subject's depressive symptoms [@pmid:27633329].
 It consists of a 30 dichotomous questionnaire, where participants are asked about their past-week feelings in reference to depressive symptoms (for instance, "Do you feel that your life is empty?"), with higher scores (i.
@@
 We used mean and standard deviation (SD) to describe continuous variables, absolute and relative frequencies to describe discrete variables.
-In order to assess the realtionship between variables, we used Pearson's product-moment correlation (r), and to assess differences between groups we used standardized mean difference (SMD) and 95% confidence intervals (CI~95%~).
+In order to assess the relationship between variables, we used Pearson's product-moment correlation (r), and to assess differences between groups we used standardized mean difference (SMD) and 95% confidence intervals (CI~95%~).
 
-As a way of controlling for the influence of psycho-physiological variables on the cardiac autonomic response of the subjects to exercise, we assess the average causal mediaton effect (ACME), the average direct effect (ADE) of the main effects after taking into account the effect of moderator variables into the observed relationships and the proportion of the effect that its mediated by these variables.
-To assess the significance of moderating variables, we used non parametric bootstrapping by using Monte Carlo draws, and bias-corrected and accelerated CI~95%~.
+As a way of controlling for the influence of psycho-physiological variables on the cardiac autonomic response of the subjects to exercise, we assess the average causal mediation effect (ACME), the average direct effect (ADE) of the main effects after taking into account the effect of moderator variables into the observed relationships and the proportion of the effect that its mediated by these variables [@kosuke2010identification].
+To assess the significance of moderating variables, we used non parametric bootstrapping by using Monte Carlo resampling, and bias-corrected and accelerated CI~95%~, using the *mediation* R package to this end [@tingley2014mediation].
 
-We defined a type I error rate at 5% ($\alpha$ = 0.05) as our threshold for null hypothesis significance testing.
+We defined a type I error rate at 5% ($\alpha$ = 0.05) as our threshold for null hypothesis significance testing, and non-overlapping CI~95%~ over the null effect for sex differences.
 All analyses were computed using the *R* programming language for statistical computing, [@rlanguage].
@@
 
-We observed that body fat percentage was linked to a lower SNS (r = -0.252, p = 0.030) and Stress index (r = 0.258, p = 0.027) response during the TSMT.
+We observed that body fat percentage was linked to a lower SNS (r = -0.252, p = 0.030) and Stress index (r = 0.258, p = 0.027) response during the intervention.
 In addition, body fat percentage was positively correlated with time domain HRV during exercise, more specifically RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020).
@@
 
+## Mediation effects associated with
+
 # Discussion
@@
 
-In addition to the potential strengths of this cross-sectional study, it is not without limitations.
-In first place, our study design prevents us from making objective inferences about the direction of the relationships, highlighting the relevance of longitudinal research exploring causal relations between our variables and underlying mechanisms.
+Despite these findings, this study is not without limitations.
+In first place, our cross-sectional design prevents us from making objective inferences about the direction of the relationships.
 In second place, a relatively small sample was included, limiting generalizability and statistical power of the analysis.

### 2024-03-01 docs/manuscript/manuscript.qmd

--- 2024-02-22 docs/manuscript/manuscript.qmd
@@
 In addition, body fat percentage was positively correlated with time domain HRV during exercise, more specifically RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020).
+No other body composition variables were found to influence the cardiac autonomic response to exercise.
 
@@
 <!--- Figura 2 --->
+
+## Mediation effects associated with
 

### 2024-03-03 docs/manuscript/manuscript.qmd
update on ms and fig 1

@@
 
-### Two-Minutes Step Test
+<!--- ESTA VARIABLE NO ES MENCIONADA EN LOS RESULTADOS, ES CUESTIONABLE SU MENCIÓN EN EL PAPER --->
+### Two-Minutes Step Test **RECONSIDERAR ESTA VARIABLE**
 
@@
 Participants are instructed to raise their right knee to a marked point on a wall as many times as possible within the given time frame, ensuring that each raise reaches at least a 70° angle at the thigh-femoral joint.
+<!--- ESTA VARIABLE NO ES MENCIONADA EN LOS RESULTADOS, ES CUESTIONABLE SU MENCIÓN EN EL PAPER --->
 
@@
 
-```{r}
+```{r prep-env}
+#| include: false
+#| message: false
+#| warning: false
+library(flextable)
+library(data.table)
+load("../../data/mindfulness.RData")
+```
+
+```{r tbl-1}
 #| echo: false
@@
 #| warning: false
-library(flextable)
 readRDS(file = "tables/tbl-descriptives.RDS") |> 
@@
 ```
+**Table 1**.
+Main sample characteristics.
+Body composition and sociodemograpic variables are displayed for the overall sample and aggregated by sex.
+Differences between males and females are displayed as SMD and CI~95%~ for continuous variables.
 
@@
 
+![](figures/fig-1.jpeg)
+**Figure 1**.
+Main correlations among psychological, cardiac autonomic and performance variables.
+Non-significant correlations (i.e., p > 0.05) were crossed with and "X".
+The more intense the color the stronger the correlation, either positive or negative.
+
 <!--- Figura con cuatro mostrando las correlaciones --->
@@
 
-In addition, a greater body fat percentage was negatively correlated with physical fitness (r = -0.273, p = 0.015).
+In addition, a greater body fat percentage was negatively correlated with the SPPB total score (r = -0.273, p = 0.015), menawhile, no other metrics of body composition were associated with physical performance of the SPPB (i.e., p > 0.05).
 
@@
 
-Physical fitness was positively correlated with a greater sympathetic activity during exercise, which was reflected by SNS index (r = 0.313, p = 0.006), mean HR (r = 0.355, p = 0.002) and its inverse counterpart mean R-R (r = -0.35, p = 0.002).
+Physical fitness, reflected through SPPB total score, was positively correlated with a greater sympathetic activity during exercise, which was reflected by SNS index (r = 0.313, p = 0.006), mean HR (r = 0.355, p = 0.002) and its inverse counterpart mean R-R (r = -0.35, p = 0.002).
 
@@
 
-## Mediation effects associated with
+## Mediation
+
+[Pendiente]
 

### 2024-03-03 R/cor_plot.R

@@ -1,0 +1,34 @@
+
+# Prepare workspace -------------------------------------------------------
+
+## Import libraries
+library(data.table)
+library(corrplot)
+
+## Import data
+data("mindfulness")
+
+## Select variables
+ind <- c(GDS = "gds_total", BAI = "bai_total", MOCA = "moca_total", HSPS = "hsps_total", SPPB = "sppb_total",
+         Height = "cc_talla", Weight = "cc_peso_corporal", BMI = "cc_imc", `Body fat` = "cc_grasa_total_porcentaje", `Muscle mass` = "cc_masa_muscular_total", 
+         Water = "cc_agua_corporal_porcentaje", `Bone mass` = "cc_masa_osea_porcentaje", 
+         `Systolic BP` = "cv_pas_inicio_mmhg", `Diastolic BP` = "cv_pad_inicio_mmhg", `Pulse pressure` = "cv_pp_media_mmhg", 
+         `Mean HR` ="hrv_mean_hr_peri", `Mean R-R` = "hrv_mean_rr_peri", RMSSD = "hrv_rmssd_peri", SDNN = "hrv_sdnn_peri",
+         LF = "hrv_lf_peri", HF = "hrv_hf_peri",  PNS = "hrv_pns_peri", SNS = "hrv_sns_peri", Stress = "hrv_stress_peri")
+
+cor_data <- mindfulness[id_event_name == "Basal", .SD, .SDcols = ind]
+names(cor_data) <- names(ind)
+
+
+corr <- cor(cor_data, use = "pairwise.complete.obs", method = "pearson")
+corr_p <- cor.mtest(cor_data, conf.level = .95, method = "pearson")
+
+fig <- corrplot(corr, type = "upper", diag = TRUE, p.mat = corr_p$p,
+                sig.level = 0.05, insig = "pch", pch = "X", pch.cex = .75,
+                tl.col = "gray20") |> substitute()
+eval(fig)
+
+jpeg("docs/manuscript/figures/fig-1.jpeg", width = 1080*2.5, height = 1080*2.5, res = 350)
+eval(fig)
+dev.off()
+

### 2024-03-03 docs/manuscript/manuscript.qmd
Reordenando ubicacion de archivos
Codigo junto a su resultado (figura)

@@
 
-```{r tbl-1}
+```{r}
 #| echo: false

### 2024-03-04 R/correlations.R
Actualización en MS, codigo y figuras
- Sección de metodos (coherencia)
- Nueva Fig 1 y respectivo caption
- Nueva seccion de mediación
- Sección de discusión (coherencia)
- Sección de conclusión (coherencia)

@@ -63,7 +63,7 @@ cor_1 <- correlation(
     id_event_name == "Basal", 
     j = .SD, 
     .SDcols = grepl(
-      pattern = "^cc|^cv|^hrv|total$", 
+      pattern = "^cc|^cv|^hrv|total$|tm_pasos", 
       x = names(mindfulness)
     )
   ],

### 2024-03-04 docs/manuscript/figures/cor_plot.R

@@ -9,12 +9,12 @@ library(corrplot)
 data("mindfulness")
 
 ## Select variables
-ind <- c(GDS = "gds_total", BAI = "bai_total", MOCA = "moca_total", HSPS = "hsps_total", SPPB = "sppb_total",
+ind <- c(`GDS-30 score` = "gds_total", `BAI score` = "bai_total", `SPPB score` = "sppb_total", `TMST steps` = "tm_pasos",
          Height = "cc_talla", Weight = "cc_peso_corporal", BMI = "cc_imc", `Body fat` = "cc_grasa_total_porcentaje", `Muscle mass` = "cc_masa_muscular_total", 
          Water = "cc_agua_corporal_porcentaje", `Bone mass` = "cc_masa_osea_porcentaje", 
          `Systolic BP` = "cv_pas_inicio_mmhg", `Diastolic BP` = "cv_pad_inicio_mmhg", `Pulse pressure` = "cv_pp_media_mmhg", 
          `Mean HR` ="hrv_mean_hr_peri", `Mean R-R` = "hrv_mean_rr_peri", RMSSD = "hrv_rmssd_peri", SDNN = "hrv_sdnn_peri",
-         LF = "hrv_lf_peri", HF = "hrv_hf_peri",  PNS = "hrv_pns_peri", SNS = "hrv_sns_peri", Stress = "hrv_stress_peri")
+         LF = "hrv_lf_peri", HF = "hrv_hf_peri",  `PNS Index` = "hrv_pns_peri", `SNS Index` = "hrv_sns_peri", `Stress Index` = "hrv_stress_peri")
 
 cor_data <- mindfulness[id_event_name == "Basal", .SD, .SDcols = ind]
 names(cor_data) <- names(ind)
@@ -28,7 +28,7 @@ fig <- corrplot(corr, type = "upper", diag = TRUE, p.mat = corr_p$p,
                 tl.col = "gray20") |> substitute()
 eval(fig)
 
-jpeg("docs/manuscript/figures/fig-1.jpeg", width = 1080*2.5, height = 1080*2.5, res = 350)
+jpeg("docs/manuscript/figures/fig-1.jpeg", width = 1080*2.5, height = 1080*2.5, res = 400)
 eval(fig)
 dev.off()
 

### 2024-03-04 docs/manuscript/manuscript.qmd

@@
 HRV was recorded through a non-invasive chest band, immediately before, during, and immediately after the execution of the two-minutes step test (TMST).
-Vital signs, including blood pressure, were monitored throughout the test, and participants' well-being was visually checked to ensure they were comfortable and mentally prepared.
-For resting HRV measurements (prior and posterior to TMST), the volunteers remained seated in a chair, while proper support of the feet and back, and avoiding talking during the recording was ensured.
-RR intervals were recorded continuously during the last 10 min of rest, and were subsequently analyzed for 5 min in each occasion.
+Vital signs, including blood pressure, were monitored throughout the test, and participants' well-being was visually checked to ensure they were comfortable and prepared.
+For resting HRV measurements (before and after the application of the TMST), the volunteers remained seated in a chair, with feet and back supported, ensuring avoiding to talk during the recordings.
+R-R intervals were recorded continuously during the last 10 min of rest, and were subsequently analyzed for 5 min in each occasion.
 The breathing rate was spontaneous.
-As part of the protocol, it was ensured that the participant had a blood pressure less than 140/90 mmHg to start the HRV measurements.
+As part of the protocol, it was ensured that the participant needed to have a blood pressure less than 140/90 mmHg to start the HRV measurements.
 
@@
 
-Physical and Physiological assessments, as well as the initial interview, were made by professional physiotherapists.
+Physical and physiological assessments, as well as the initial interview, were made by professional physiotherapists.
 
@@
 
-Cardiac autonomic activity was assessed using RR interval recordings (HRV) obtained through the Polar Team2 system (Polar®) application.
+Cardiac autonomic activity was assessed using R-R interval recordings (HRV) obtained through the Polar Team2 system (Polar®) application.
 Artifacts and ectopic heartbeats, which did not exceed 3% of the recorded data, were excluded [@malik1996heart].
-Time-domain parameters considered for analysis included the square root of the mean squared differences of successive RR intervals (RMSSD, expressed in ms) as an index of parasympathetic activity [@pmid:20033207], and the standard deviation of RR intervals (SDNN), reflecting total variability encompassing both sympathetic and parasympathetic contributions to cardiac autonomic function [@pmid:9401419; @buchheit2006cardiac].
-Additionally, the Stress Index (SI) and Parasympathetic and Sympathetic Nervous System Index (PNS and SNS, respectively) were calculated.
+Time-domain parameters considered for analysis included the square root of the mean squared differences of successive R-R intervals (RMSSD, expressed in ms) as an index of parasympathetic activity [@pmid:20033207], and the standard deviation of RR intervals (SDNN), reflecting total variability encompassing both sympathetic and parasympathetic contributions to cardiac autonomic function [@pmid:9401419; @buchheit2006cardiac].
+Additionally, the Stress Index (SI) and Parasympathetic and Sympathetic Nervous System Index (PNS and SNS, respectively) were computed.
 The PNS Index, indicative of total vagal stimulation, was derived from mean R-R intervals, RMSSD, and Poincaré Plot Index SD1 in normalized units (linked to RMSSD), representing deviations from normal population averages [@pmid:9401419; @GSID:XkAQwTGLgNsJ].
@@
 The SI serves as an indicator representing the ANS control system's workload [@doi:10.1159/000513781], normalized by the square root of Baevsky’s SI [@2009MethodicalRU].
-All analyses were conducted using Kubios HRV® software (Kuopio, Finland).
+All analyses conducted to compute HRV related indices were performed using Kubios HRV® software (Kuopio, Finland).
 
-<!--- ESTA VARIABLE NO ES MENCIONADA EN LOS RESULTADOS, ES CUESTIONABLE SU MENCIÓN EN EL PAPER --->
-### Two-Minutes Step Test **RECONSIDERAR ESTA VARIABLE**
+### Two-Minutes Step Test
 
@@
 Participants are instructed to raise their right knee to a marked point on a wall as many times as possible within the given time frame, ensuring that each raise reaches at least a 70° angle at the thigh-femoral joint.
-<!--- ESTA VARIABLE NO ES MENCIONADA EN LOS RESULTADOS, ES CUESTIONABLE SU MENCIÓN EN EL PAPER --->
+The number of valid steps were recorded for each subject.
 
@@
 As a way of controlling for the influence of psycho-physiological variables on the cardiac autonomic response of the subjects to exercise, we assess the average causal mediation effect (ACME), the average direct effect (ADE) of the main effects after taking into account the effect of moderator variables into the observed relationships and the proportion of the effect that its mediated by these variables [@kosuke2010identification].
-To assess the significance of moderating variables, we used non parametric bootstrapping by using Monte Carlo resampling, and bias-corrected and accelerated CI~95%~, using the *mediation* R package to this end [@tingley2014mediation].
+To assess the significance of moderating variables, we used non overlapping CI~95%~, estimated based on non parametric bootstrapping by using Monte Carlo resampling, and bias-corrected and accelerated CI~95%~, using the *mediation* R package to this end [@tingley2014mediation].
 
-We defined a type I error rate at 5% ($\alpha$ = 0.05) as our threshold for null hypothesis significance testing, and non-overlapping CI~95%~ over the null effect for sex differences.
+We defined a type I error rate at 5% ($\alpha$ = 0.05) as our threshold for null hypothesis significance testing, and non-overlapping CI~95%~ over the null effect for sex differences and mediation analyses.
 All analyses were computed using the *R* programming language for statistical computing, [@rlanguage].
@@
 
-<!---
-#> cc_grasa_total_porcentaje, hrv_rmssd_peri, r = 0.253, p = 0.029
-#> cc_grasa_total_porcentaje, hrv_rmssd_peri, rho = 0.249, p = 0.032
-#> 
-#> cc_grasa_total_porcentaje, hrv_stress_peri, r = -0.258, p = 0.027
-#> cc_grasa_total_porcentaje, hrv_stress_peri, rho = -0.282, p = 0.015
-#> 
-#> cc_grasa_total_porcentaje, hrv_sns_peri, r = -0.252, p = 0.03
-#> cc_grasa_total_porcentaje, hrv_sns_peri, rho = -0.257, p = 0.027
-#> 
-#> cc_grasa_total_porcentaje, hrv_sdnn_peri, r = 0.269, p = 0.02
-#> cc_grasa_total_porcentaje, hrv_sdnn_peri, rho = 0.284, p = 0.014
---->
-
 ![](figures/fig-1.jpeg)
@@
 The more intense the color the stronger the correlation, either positive or negative.
-
-<!--- Figura con cuatro mostrando las correlaciones --->
+GDS, Geriatric Depression Scale; BAI, Beck Anxiety Inventory; SPPB, Short Physical Performance Battery; TMST, Two-Minutes Step Test; BMI, Body Mass Index.
 
@@
 
-In addition, a greater body fat percentage was negatively correlated with the SPPB total score (r = -0.273, p = 0.015), menawhile, no other metrics of body composition were associated with physical performance of the SPPB (i.e., p > 0.05).
-
-<!---
-#> cc_grasa_total_porcentaje, sppb_total, r = -0.273, p = 0.015
-#> cc_grasa_total_porcentaje, sppb_total, rho = -0.25, p = 0.026
---->
+In addition, greater SPPB score was observed in leaner (r = -0.273, p = 0.015) and individuals with lower BMI values (r = -0.225, p = 0.047).
+Similar findings were observed between total steps from the TMST and body fat percentage (r = -0.279, p = 0.013) and bodyweight (r = -0.232, p = 0.041).
+Not enough evidence was observed to suggest a correlation between total muscle mass with either total SPPB score (r = -0.063, p = 0.584) or total number of steps in TMST (r = -0.070, p = 0.541).
 
@@
 
-Physical fitness, reflected through SPPB total score, was positively correlated with a greater sympathetic activity during exercise, which was reflected by SNS index (r = 0.313, p = 0.006), mean HR (r = 0.355, p = 0.002) and its inverse counterpart mean R-R (r = -0.35, p = 0.002).
+Physical fitness, reflected through SPPB total score and total steps in TMST, was positively correlated with a greater sympathetic activity during exercise, which was reflected by SNS index (SPPB score, r = 0.313, p = 0.006; TMST steps, r = 0.265, p = 0.022), mean HR (SPPB score, r = 0.355, p = 0.002; TMST steps, r = 0.338, p = 0.003) and its inverse counterpart mean R-R (SPPB score, r = -0.35, p = 0.002; TMST steps, r = -0.311, p = 0.007).
+Total number of steps from TMST was also inversely correlated with parasympathetic indices during exercise like RMSSD (r = -0.285, p = 0.014) and PNS index (r = -0.344, p = 0.003).
 
-<!--- 
-#> hrv_sns_peri, sppb_total, r = 0.313, p = 0.006
-#> hrv_sns_peri, sppb_total, rho = 0.332, p = 0.004
+However, this appears to be accompanied by an proportional decrease in parasympathetic drive during exercise with increasing levels of physical fitness displayed in the SPPB, which was denoted by the inverse correlation of the later with PNS index during (r = -0.375, p = 0.001) and after exercise (r = -0.229, p = 0.049), in addition to RMSSD during exercise (r = -0.294, p = 0.010)
 
-#> hrv_mean_hr_peri, sppb_total, r = 0.355, p = 0.002
-#> hrv_mean_hr_peri, sppb_total, rho = 0.375, p = 0.001
+## Mediation analysis
 
-#> hrv_mean_rr_peri, sppb_total, r = -0.35, p = 0.002
-#> hrv_mean_rr_peri, sppb_total, rho = -0.38, p = 0.001
+Mediation analyses suggest many potential influential effects for the relationships for body composition and parasympathetic indices.
+In this context, the average direct effect (ADE) of body fat percentage on RMSSD during exercise (ADE = 0.088, CI~95%~[-0.005, 0.200], p = 0.077) was accentuated when considering the influence of SPPB score (ACME = 0.0289, CI~95%~[0.003, 0.110], p = 0.057; Total effect = 0.117, CI~95%~[0.028, 0.240], p = 0.012).
+Similar findings were observed when considering the mediation effect of TMST steps into account (ACME = 0.0283, CI~95%~[0, 0.07], p = 0.079; Total effect = 0.117, CI~95%~[0.028, 0.240], p = 0.012).
+No other parasympathetic indicators were influenced or mediated by fitness or psychological related measures.
 
---->
+Additionally, and in the case on sympathetic indicators, the effect of body fat on SNS index during exercise (ADE = -0.050, CI~95%~[-0.107, 0.01], p = 0.077) was influenced by SPPB score in similar way as with RMSSD, potentiating the original main observed effect (ACME = -0.016, CI~95%~[-0.049, 0], p = 0.041; Total effect = -0.066, CI~95%~[-0.125, -0.010], p = 0.016).
+Similar effects were observed with TMST steps in this regard (ACME = -0.015, CI~95%~[-0.042, 0], p = 0.050; Total effect = -0.066, CI~95%~[-0.125, -0.010], p = 0.016).
+No other sympathetic indices were notoriously modified in the presence of either SPPB score, TMST steps or psychological variables.
 
-However, this appears to be accompanied by an proportional decrease in parasympathetic drive during exercise with increasing levels of physical fitness, which was denoted by the inverse correlation of the later with PNS index during (r = -0.375, p = 0.001) and after exercise (r = -0.229, p = 0.049), in addition to RMSSD during exercise (r = -0.294, p = 0.010)
-
-<!--- 
-#> hrv_pns_peri, sppb_total, r = -0.375, p = 0.001
-#> hrv_pns_peri, sppb_total, rho = -0.372, p = 0.001
-
-#> hrv_pns_post, sppb_total, r = -0.229, p = 0.049
-#> hrv_pns_post, sppb_total, rho = -0.238, p = 0.04
-
-#> hrv_rmssd_peri, sppb_total, r = -0.294, p = 0.01
-#> hrv_rmssd_peri, sppb_total, rho = -0.282, p = 0.014
---->
-
-<!--- Figura 2 --->
-
-## Mediation
-
-[Pendiente]
+When assessing potential mediators for the effect of body composition on fitness related measures, we found that the effect of body fat percentage on SPPB score (ADE = -0.075, CI~95%~[-0.123, -0.020], p = 0.007) is partially diminished when considering the influence of total muscle mass (ACME = 0.011, CI~95%~[0, 0.040], p = 0.107; Total effect = -0.064, CI~95%~[-0.111, -0.010], p = 0.014).
+Similar mediation effects of total muscle mass were observed (ACME = 0.118, CI~95%~[0.005, 0.350], p = 0.075) when assessing the effect of body fat percentage on TMST steps (ADE = -0.773, CI~95%~[-1.193, -0.300], p = 0.001; Total effect = -0.655, CI~95%~[-1.084, -0.210], p = 0.004).
+No other mediating effects were observed for any of the psychological measures for any of the relationships between body composition and physical fitness measures.
 
@@
 This study aimed to investigate the relationship between physical fitness, body fat percentage, and cardiac autonomic response to exercise in older people.
-Through the correlational analysis, we aimed to elucidate the  interplay between these factors and their implications for cardiac autonomic function during physical efforts.
+Through correlational analyses, we aimed to elucidate the  interplay between these factors and their implications for cardiac autonomic function during physical efforts.
 
-It was hypothesized that a higher body fat percentage would be inversely related to  physical aptitude.
-Our results aligns with this hypothesis and previous research regarding body composition changes and physical performance decrease during aging [@pmid:35832415; @pmid:36256894].
+It was hypothesized that a higher body fat percentage would be inversely correlated to physical fitness.
+Our results aligns with this hypothesis and with those of previous research regarding body composition and physical performance, with a proportional decrease in aging individuals [@pmid:35832415; @pmid:36256894].
 
-Furthermore, the increase in body fat percentage was related to a reduced sympathetic activity response and a higher parasympathetic activity during TMST, as hypothesized (i.
-e., SNS index, SI, SDNN and RMSSD).
-Throughout the course of physical efforts, ANS is expected to shift into a sympathetic predominance, ensuring a sufficient response to higher metabolic demands [@pmid:33246163].
+Furthermore, greater body fat percentage was related to a reduced sympathetic drive and higher parasympathetic activity during the TMST, as hypothesized.
+Throughout the course of physical efforts, the ANS is expected to shift into a sympathetic state, ensuring an efficient response to greater metabolic demands [@pmid:33246163].
 However, aged individuals may present an impaired ANS response to physical exercise [@ijerph20054456].
-In that sense, our results support that body composition may play a role in that response, opening opportunities for targeted interventions.
+In that sense, our results support that body composition may play a pivotal role in that response, generating novel insights for potential interventions.
 
-Nonetheless, the moderating effect of physical fitness on the relationship proposed above, suggests that more physically active individuals may counteract the negative effect of a high body fat percentage on cardiac autonomic response to exercise.
+Nonetheless, the moderating effect of physical fitness related measures on the relationship proposed earlier, suggests that greater physical activity could potentially counteract the negative effects of a high body fat percentage on cardiac autonomic response to exercise, possibly, at the same extent as seen in leaner individuals.
 This result highlights the importance of promoting a physically active lifestyle and favor the maintenance of physical capacity, regardless of nutritional status [@pmid:33402165; @pmid:30002462].
 
-Regarding physical fitness and cardiac autonomic response to exercise, as expected, our results evidenced that individuals with a higher sympathetic activity during exercise, as measured by SNS index, had a greater physical fitness while the opposite observation was made for parasympathetic activity (i.
-e., PNS index and RMSSD).
-Previous research had demonstrated the utility of HRV decrease as an index of cardiorespiratory fitness [@pmid:36054204], supporting the hypothesis that a higher sympathetic activity facilitates the performance during physical moderate effort.
-Even more, it has been previously demonstrated that physically active individuals may have an enhanced autonomic response to exercise, strengthening this relationship [@pmid:28721030; @pmid:24799513].
+Regarding the link between physical fitness and cardiac autonomic response to exercise, our results evidenced that individuals with a high sympathetic drive while exercising, had greater physical fitness.
+Additionally, previous research had demonstrated the utility of HRV and exercise-induced responses in those metrics as a potential marker of cardiorespiratory fitness in these individuals [@pmid:36054204], supporting the current hypothesis that a greater sympathetic drive facilitates performance and its related measures during moderate intensity physical activities.
+Even more, it has been previously demonstrated that physically active individuals have an enhanced autonomic response to exercise, playing a pivotal role strengthening this hypothesis [@pmid:28721030; @pmid:24799513].
 
-Our results demonstrate the interplay between physical fitness, body fat, and cardiac autonomic response to exercise in older individuals.
-Physical fitness is negatively affected by higher body fat percentage , while also being associated with impaired cardiac autonomic response to exercise, characterized by decreased sympathetic activity and increased parasympathetic tone.
-Additionally, this altered autonomic response is related to lower physical performance in older adults, both variables being important predictors of health outcomes in this population [@pmid:30627571; @pmid:30104034].
-Overall, our findings underscore the triangular relationship among these variables.
+Our results provide evidence to support the interplay between physical fitness, body fat, and cardiac autonomic response to exercise in older individuals.
+Physical fitness is negatively affected by higher body fat percentage, while also being associated with impaired cardiac autonomic response to exercise, characterized by decreased sympathetic drive and increased parasympathetic tone.
+Additionally, this altered autonomic response is related to lower physical performance in older individuals, both variables being important predictors of health outcomes in this population [@pmid:30627571; @pmid:30104034].
+Overall, our findings underscore the triangular and bidirectional relationship among these variables.
 
-However, our results did not evidence a moderating effect of GDS-30 and BAI scores, challenging initial hypothesis.
-This finding suggests a nuanced interplay between psychological variables and physical/physiological measures in older people, underscoring the need for further research to elucidate underlying mechanisms and variables beyond depression and anxiety.
-While the lack of moderating effects does not diminish the significance of psychological factors in aging, it highlights the complexity of their influence.
+However, our results did not evidence a moderating effect of either GDS-30 or BAI scores, challenging our initial hypothesis.
+These finding suggests a nuanced interplay between psychological variables and physical/physiological measures in older people, underscoring the need for further research to elucidate underlying mechanisms and variables beyond depression and anxiety.
+While the lack of evidence suggesting any moderating effects does not diminish the significance of psychological factors in aging, it highlights the complexity of their influence.
 
 In addition to the potential strengths of this cross-sectional study, it is not without limitations.
-In first place, our study design prevents us from making objective inferences about the direction of the relationships, highlighting the relevance of longitudinal research exploring causal relations between our variables and underlying mechanisms.
-In second place, a relatively small sample was included, limiting generalizability and statistical power of the analysis.
+First, our study design prevents us from making causal inferences about the direction of the relationships, highlighting the relevance of experimental and longitudinal research exploring causal relations between our variables and potential underlying mechanisms.
+Second, a relatively small sample was included, limiting generalizability and statistical power of the analysis.
 Furthermore, the sample consisted mainly of women (82.7%), limiting our ability to make sex comparisons and decreasing the representativeness of the sample.
@@
 
-The results highlighted a triangular relationship between the variables.
-Body fat percentage was importantly related to both physical fitness and cardiac autonomic activity during exercise, while sympathetic activity was associated with a grater physical fitness in older people.
-Furthermore, physical fitness may moderate the associations between body fat and cardiac autonomic activity, highlighting its importance in older people's health.
+The results highlighted a structural bidirectional relationship among variables.
+Body fat percentage was inversely correlated to both physical fitness and cardiac autonomic activity during exercise, with sympathetic drive associated with physical fitness in older people.
+Furthermore, physical fitness may mediate the effect of body fat on cardiac autonomic activity during exercise, highlighting its pivotal role in older people's health.
 However, further research is needed to elucidate moderating effects of psychological variables in this field.

### 2024-03-08 docs/manuscript/manuscript.qmd
Ms minor Changes
update in discussion - mediating effects discussed - strenghts and limitations updated - some citations are still required [cita requerida]

@@
 Also, a negative link existed between body fat and sympathetic activity: SNS index (r = -0.252, p = 0.030) and Stress index (r = 0.258, p = 0.027) during TMST.
-But a positive link with parasympathetic activity: RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020) **Conclusion**: the findings unveil a triangular relationship, where body fat percentage significantly impacts physical fitness and cardiac autonomic activity during exercise, while a greater sympathetic activity correlates with higher physical fitness.
-Additionally, physical fitness moderates the link between body fat and cardiac autonomic activity, emphasizing its importance for older adults' health.
-Further research is needed to explore psychological variables' moderating effects.
+But a positive link with parasympathetic activity: RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020) **Conclusion**: Body fat was inversely correlated to both physical fitness and cardiac autonomic activity during exercise, with sympathetic drive associated with physical fitness in older people.
+Furthermore, physical fitness may mediate the effect of body fat on cardiac autonomic activity during exercise, and total muscle mass may mitigate the negative effect of increased adiposity on physical fitness, highlighting its pivotal role in older people's health.
+However, further research is needed to elucidate mediating effects of psychological variables in this field.
 
@@
 Physical fitness has been classically defined as the ability to perform daily tasks with vigor and without undue fatigue, and with ample energy to enjoy leisure-time pursuits and to meet unforeseen emergencies [@pmid:3920716].
-Therefor, it is a crucial component of healthy aging.
+Therefore, it is a crucial component of healthy aging.
 The term encompasses a multitude of aspects, typically including cardiorespiratory endurance, muscular endurance and strength, body composition, and flexibility [@pmid:3920711].
-Numerous studies have demonstrated the positive impact of physical fitness on overall health and longevity, highlighting its role in reducing the risk of chronic diseases, and improving quality of life and functional independence in older adults [@pmid:9177581; @pmid:29293447; @pmid:30768197; @pmid:36332759]
+Numerous studies have demonstrated the positive impact of physical fitness on overall health and longevity, emphasizing its role in reducing the risk of chronic diseases, and improving quality of life and functional independence in older adults [@pmid:9177581; @pmid:29293447; @pmid:30768197; @pmid:36332759]
 
@@
 
-Despite extensive research into physical fitness, body composition, and cardiovascular health in older adults, gaps persist in the understanding of their interrelationships and underlying mechanisms.
+Despite extensive research into physical fitness, body composition, and cardiovascular health in older adults, gaps persist in the understanding of their interrelationships and factors mediating them.
 One area of interest is the relationship between physical fitness, body fat percentage, and cardiac autonomic response to exercise in older individuals.
-Although previous studies have explored these factors independently, this study explore their collective influence on physical fitness in the aging population.
+This study explores their collective influence on physical fitness in the aging population.
 
-The conceptualization of this study is based on the recognition of the interplay between physical fitness, body composition, and autonomic cardiovascular regulation in aging.
+The conceptualization of this investigation is based on the recognition of the interplay between physical fitness, body composition, and autonomic cardiovascular regulation in aging.
 We aim to answer the following research question: How does physical fitness relate to body fat percentage and cardiac autonomic response to exercise in older people?
 By elucidating these correlations, we hope to gain valuable insights into the relationships that underlie age-related modifications in physical fitness and identify potential interventions to enhance health outcomes in older adults.
-Even more, we seek to understand the moderating effect of frequent psychological traits, such as geriatric depression and anxiety.
+Even more, we seek to understand the mediating effect of the variables, including frequent psychological traits, such as geriatric depression and anxiety.
 
@@
 The primary objective of this study is to investigate the relationship between physical fitness, body fat percentage, and cardiac autonomic response to exercise in older people.
-Even more, We aimed to assess the moderating effect of psychological variables on these relationships, with implications for cardiovascular health and functional capacity in aging.
+Even more, We aimed to assess the mediating effect of psychological variables on these relationships, with implications for cardiovascular health and functional capacity in aging.
 
@@
 81 community-dwelling older adults were recruited and selected by non-probabilistic sampling, from the Region of Magallanes and Chilean Antartica, Chile.
-They were included if: 65 years age or older; residents of the Region of Magallanes and Chilean Antartica, Chile; and being able to understand the study aims and instructions.
+They were included if: 65 years age or older; residents of the Region of Magallanes and Chilean Antarctica, Chile; and being able to understand the study aims and instructions.
 But they were excluded in case of: diagnosis of congenital heart disease; consumption of beta-blocker drugs, taking stimulant substances within 24 h before assessments session; motor or cognitive disability; inability to understand instructions or written content; or presence of pain during cardiac or physical assessments.
@@
 
-Additionally, and in the case on sympathetic indicators, the effect of body fat on SNS index during exercise (ADE = -0.050, CI~95%~[-0.107, 0.01], p = 0.077) was influenced by SPPB score in similar way as with RMSSD, potentiating the original main observed effect (ACME = -0.016, CI~95%~[-0.049, 0], p = 0.041; Total effect = -0.066, CI~95%~[-0.125, -0.010], p = 0.016).
+Additionally, and in the case on sympathetic indicators, the effect of body fat on SNS index during exercise (ADE = -0.050, CI~95%~[-0.107, 0.01], p = 0.077) was influenced by SPPB score in similar way as with RMSSD, enhancing the original main observed effect (ACME = -0.016, CI~95%~[-0.049, 0], p = 0.041; Total effect = -0.066, CI~95%~[-0.125, -0.010], p = 0.016).
 Similar effects were observed with TMST steps in this regard (ACME = -0.015, CI~95%~[-0.042, 0], p = 0.050; Total effect = -0.066, CI~95%~[-0.125, -0.010], p = 0.016).
@@
 
+Main results are represented in the Figure X.
+
+Fig. X. Diagram of main results.
+
+****Insertar Figura resumen resultados****
+
+Footnote: The figure represents the main results of this study.
+Blue arrows and boxes represent the correlation between variables, while orange arrows and boxes represent the factors mediating the above correlations.
+"+" means that the mediating factor enhances the main observed effect, contrary to "-", reflecting a diminished main effect when considering the mediating factor.
+
 # Discussion
@@
 This study aimed to investigate the relationship between physical fitness, body fat percentage, and cardiac autonomic response to exercise in older people.
-Through correlational analyses, we aimed to elucidate the  interplay between these factors and their implications for cardiac autonomic function during physical efforts.
+Through correlational and mediating analyses, we aimed to elucidate the interplay between these factors and their implications for cardiac autonomic function during physical efforts.
 
@@
 
-Furthermore, greater body fat percentage was related to a reduced sympathetic drive and higher parasympathetic activity during the TMST, as hypothesized.
-Throughout the course of physical efforts, the ANS is expected to shift into a sympathetic state, ensuring an efficient response to greater metabolic demands [@pmid:33246163].
-However, aged individuals may present an impaired ANS response to physical exercise [@ijerph20054456].
-In that sense, our results support that body composition may play a pivotal role in that response, generating novel insights for potential interventions.
-
-Nonetheless, the moderating effect of physical fitness related measures on the relationship proposed earlier, suggests that greater physical activity could potentially counteract the negative effects of a high body fat percentage on cardiac autonomic response to exercise, possibly, at the same extent as seen in leaner individuals.
-This result highlights the importance of promoting a physically active lifestyle and favor the maintenance of physical capacity, regardless of nutritional status [@pmid:33402165; @pmid:30002462].
+However, the mediating analysis showed that total muscle mass diminished the main observed effect of body fat on physical fitness, with more muscled individuals displaying a better physical performance despite the body fat percentage.
+This finding suggests that older adults could be benefited of activities that promotes their muscle mass conservation, even when they have high body fat percentages [cita requerida].
+Nevertheless, caution is advised in the interpretation of this result, as the close physiological association among increasing body fat and decreasing muscle mass during aging have been well-studied [cita requerida: sarcopenic obesity]
 
 Regarding the link between physical fitness and cardiac autonomic response to exercise, our results evidenced that individuals with a high sympathetic drive while exercising, had greater physical fitness.
-Additionally, previous research had demonstrated the utility of HRV and exercise-induced responses in those metrics as a potential marker of cardiorespiratory fitness in these individuals [@pmid:36054204], supporting the current hypothesis that a greater sympathetic drive facilitates performance and its related measures during moderate intensity physical activities.
-Even more, it has been previously demonstrated that physically active individuals have an enhanced autonomic response to exercise, playing a pivotal role strengthening this hypothesis [@pmid:28721030; @pmid:24799513].
+Previous research had demonstrated the utility of HRV and exercise-induced responses in those metrics as a potential marker of cardiorespiratory fitness in these individuals [@pmid:36054204], supporting the current hypothesis that a greater sympathetic drive facilitates physical performance during moderate intensity activities.
+Even more, it has been previously demonstrated that physically active individuals have an enhanced autonomic response to exercise, playing a pivotal role and strengthening this hypothesis [@pmid:28721030; @pmid:24799513].
+
+Furthermore, the increase in body fat percentage was related to a reduced sympathetic activity and a higher parasympathetic activity during TMST, as hypothesized (i.e., SNS index, SI, SDNN and RMSSD).
+Throughout the course of physical efforts, ANS is expected to shift into a sympathetic predominance, ensuring a sufficient response to higher metabolic demands [@pmid:33246163].
+However, aged individuals may present an impaired ANS response to physical exercise [@ijerph20054456].
+In that sense, our results support that body composition plays a role in autonomic response.
+Moreover, they suggest that the negative effects of body fat on physical fitness in older adults, may be partially attributable to less efficient autonomic inputs to the cardiovascular system [cita requerida].
+
+Surprisingly, mediating analyses found that higher physical fitness (SPPB and TMST scores) intensifies the negative impact of body fat on autonomic control.
+This unexpected finding hints at a complex interplay between physical fitness, adiposity, and cardiac autonomic regulation.
+Our results indicate that in individuals with greater functional capacity, the influence of body fat on HRV is amplified rather than mitigated.
+This observation challenges traditional conceptions about the relationship between physical fitness and cardiac autonomic regulation, and raises questions about the underlying mechanisms involved.
+
+As far as we know, there are not preceding studies addressing these findings.
+We posit that there exists further factors influencing the mediation effect.
+For instance, body distribution of the adipose tissue was not explored, specially when it is known that visceral fat has an important impact on cardiovascular health and its autonomic regulation [cita requerida].
+Participants with higher fitness may have a different fat distribution compared with lower performance individuals, which may induce different effects on autonomic drive to the heart [cita requerida].
+Furthermore, other physical physiological, demographic or psychological variables may be playing an unexplored role.
+However, these hypothesis must be examined under scientific standards.
+Therefore, exhaustive research is needed to address this intriguing interplay.
 
@@
 
-However, our results did not evidence a moderating effect of either GDS-30 or BAI scores, challenging our initial hypothesis.
+However, our results did not evidence a mediating effect of either GDS-30 or BAI scores, challenging our initial hypothesis.
 These finding suggests a nuanced interplay between psychological variables and physical/physiological measures in older people, underscoring the need for further research to elucidate underlying mechanisms and variables beyond depression and anxiety.
-While the lack of evidence suggesting any moderating effects does not diminish the significance of psychological factors in aging, it highlights the complexity of their influence.
+While the lack of evidence suggesting any mediating effects does not diminish the significance of psychological factors in aging, it highlights the complexity of their influence.
 
-In addition to the potential strengths of this cross-sectional study, it is not without limitations.
+This study was strengthened by the fact that both SPPB and TMST measures consistently aligned in their interaction with other variables, which supports a global understanding of physical fitness in older people, rather than specific aspects, and by the exploration of mediating factors in the primary results.
+Nonetheless, this cross-sectional study, it is not without limitations.
 First, our study design prevents us from making causal inferences about the direction of the relationships, highlighting the relevance of experimental and longitudinal research exploring causal relations between our variables and potential underlying mechanisms.
@@
 Furthermore, the sample consisted mainly of women (82.7%), limiting our ability to make sex comparisons and decreasing the representativeness of the sample.
-It should also be noted that the relationships between physical fitness and cardiac autonomic response to exercise were not made within the same instrument, due to the possible lack of SPPB to generate an adequate physical stress, and of TMST to inform general physical fitness of the participants.
 Further research could investigate the impact of cardiac autonomic response on performance during the same task, in this age group.
@@
 Body fat percentage was inversely correlated to both physical fitness and cardiac autonomic activity during exercise, with sympathetic drive associated with physical fitness in older people.
-Furthermore, physical fitness may mediate the effect of body fat on cardiac autonomic activity during exercise, highlighting its pivotal role in older people's health.
-However, further research is needed to elucidate moderating effects of psychological variables in this field.
+Furthermore, physical fitness may mediate the effect of body fat on cardiac autonomic activity during exercise, and total muscle mass may mitigate the negative effect of increased adiposity on physical fitness, highlighting its pivotal role in older people's health.
+However, further research is needed to elucidate mediating effects of psychological variables in this field.
 

### 2024-03-08 docs/manuscript/manuscript.qmd
added refrences to discussion

@@
 However, the mediating analysis showed that total muscle mass diminished the main observed effect of body fat on physical fitness, with more muscled individuals displaying a better physical performance despite the body fat percentage.
-This finding suggests that older adults could be benefited of activities that promotes their muscle mass conservation, even when they have high body fat percentages [cita requerida].
-Nevertheless, caution is advised in the interpretation of this result, as the close physiological association among increasing body fat and decreasing muscle mass during aging have been well-studied [cita requerida: sarcopenic obesity]
+This finding suggests that older adults could be benefited of activities that promotes their muscle mass conservation, even when they have high body fat percentages [pmid:27663845].
+Nevertheless, caution is advised in the interpretation of this result, as the close physiological association among increasing body fat and decreasing muscle mass during aging have been well-studied [@pmid:33335909]
 
@@
 In that sense, our results support that body composition plays a role in autonomic response.
-Moreover, they suggest that the negative effects of body fat on physical fitness in older adults, may be partially attributable to less efficient autonomic inputs to the cardiovascular system [cita requerida].
+Moreover, they suggest that the negative effects of body fat on physical fitness in older adults, may be partially attributable to less efficient autonomic inputs to the cardiovascular system [@pmid:36875496].
 
@@
 We posit that there exists further factors influencing the mediation effect.
-For instance, body distribution of the adipose tissue was not explored, specially when it is known that visceral fat has an important impact on cardiovascular health and its autonomic regulation [cita requerida].
-Participants with higher fitness may have a different fat distribution compared with lower performance individuals, which may induce different effects on autonomic drive to the heart [cita requerida].
+For instance, body distribution of the adipose tissue was not explored, specially when it is known that visceral fat has an important impact on cardiovascular health and its autonomic regulation [@pmid:31553779; pmid:32158768].
+Participants with higher fitness may have a different fat distribution compared with lower performance individuals, which may induce different effects on autonomic drive to the heart [@pmid:19752873].
 Furthermore, other physical physiological, demographic or psychological variables may be playing an unexplored role.

### 2024-03-09 docs/manuscript/manuscript.qmd
figuras de dispersión

--- 2024-03-08 docs/manuscript/manuscript.qmd
@@
 library(data.table)
+library(ggplot2)
 load("../../data/mindfulness.RData")
+dat <- mindfulness[id_event_name == "Basal", -2L]
 ```
@@
 
-We observed that body fat percentage was linked to a lower SNS (r = -0.252, p = 0.030) and Stress index (r = 0.258, p = 0.027) response during the TSMT.
+We observed that body fat percentage was linked to a lower SNS (r = -0.252, p = 0.030) and Stress index (r = -0.258, p = 0.027) response during the TSMT.
 In addition, body fat percentage was positively correlated with time domain HRV during exercise, more specifically RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020).
@@
 
-![](figures/fig-1.jpeg)
-**Figure 1**.
-Main correlations among psychological, cardiac autonomic and performance variables.
-Non-significant correlations (i.e., p > 0.05) were crossed with and "X".
-The more intense the color the stronger the correlation, either positive or negative.
-GDS, Geriatric Depression Scale; BAI, Beck Anxiety Inventory; SPPB, Short Physical Performance Battery; TMST, Two-Minutes Step Test; BMI, Body Mass Index.
+```{r}
+#| echo: false
+#| warning: false
+#| error: false
+#| fig-width: 8
+#| fig-height: 4
+#| fig-dpi: 300
+dat_long <- melt.data.table(
+  data = dat,
+  id.vars = c("id_record", "cc_grasa_total_porcentaje"), 
+  measure.vars = c("hrv_sns_peri", "hrv_stress_peri", "hrv_rmssd_peri", "hrv_sdnn_peri", "hrv_pns_peri")
+)
+
+dat_long[, variable := fcase(
+  grepl("sns", variable), "SNS index",
+  grepl("pns", variable), "PNS index",
+  grepl("stress", variable), "Stress index",
+  grepl("sdnn", variable), "SDNN",
+  grepl("rmssd", variable), "RMSSD"
+)]
+
+dat_long[, variable := factor(variable, levels = c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index"))]
+
+ggplot(dat_long, aes(value, cc_grasa_total_porcentaje)) +
+  facet_grid(~ variable, scales = "free_x") +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  geom_text(mapping = aes(label = text), data = data.frame(
+    variable = as.factor(c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index")),
+    cc_grasa_total_porcentaje = c(61, 61, 61, 61, 61),
+    value = c(13, 20, -2, 8, 45),
+    text = c("p = 0.029", "p = 0.020", "p = 0.262", "p = 0.030", "p = 0.026")
+  ), size = 3.5, fontface = "italic") +
+  labs(y = "Body fat (%)", x = "Values") +
+  tidybayes::theme_ggdist()
+```
+
+**Figure 1**. Bivariate dispersion plots between body fat and HRV related measures. Significance values for the Pearson's product-moment correlation test is shown.
 
@@
 
+```{r}
+#| echo: false
+#| warning: false
+#| error: false
+#| fig-width: 8
+#| fig-height: 4
+#| fig-dpi: 300
+dat_long <- melt.data.table(
+  data = dat,
+  id.vars = c("id_record", "cc_grasa_total_porcentaje", "cc_masa_muscular_total"), 
+  measure.vars = c("tm_pasos", "sppb_total")
+)
+
+dat_long[, variable := fcase(
+  grepl("tm_pasos", variable), "TMST steps",
+  grepl("sppb_total", variable), "SPPB total score"
+)]
+
+dat_long[, variable := factor(variable, levels = c("TMST steps", "SPPB total score"))]
+
+fig2a <- ggplot(dat_long, aes(value, cc_grasa_total_porcentaje)) +
+  facet_grid(~ variable, scales = "free_x") +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  geom_text(mapping = aes(label = text), data = data.frame(
+    variable = as.factor(c("TMST steps", "SPPB total score")),
+    cc_grasa_total_porcentaje = c(60, 60),
+    value = c(42, 4),
+    text = c("p = 0.013", "p = 0.015")
+  ), size = 3.5, fontface = "italic") +
+  labs(y = "Body fat (%)", x = NULL) +
+  scale_y_continuous(expand = c(.1,.1)) +
+  tidybayes::theme_ggdist()
+
+fig2b <- ggplot(dat_long, aes(value, cc_masa_muscular_total)) +
+  facet_grid(~ variable, scales = "free_x") +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  geom_text(mapping = aes(label = text), data = data.frame(
+    variable = as.factor(c("TMST steps", "SPPB total score")),
+    cc_masa_muscular_total = c(65.5, 65.5),
+    value = c(42, 4),
+    text = c("p = 0.541", "p = 0.584")
+  ), size = 3.5, fontface = "italic") +
+  labs(y = "Muscle mass (kg)", x = NULL) +
+  scale_y_continuous(expand = c(.1,.1)) +
+  tidybayes::theme_ggdist()
+
+ggpubr::ggarrange(
+  fig2a, fig2b, nrow = 1
+)
+```
+**Figure 2**. Bivariate dispersion plots between physical fitness and body composition related measures. Significance values for the Pearson's product-moment correlation test is shown.
+
 ## Physical fitness and cardiac autonomic response to exercise
@@
 However, this appears to be accompanied by an proportional decrease in parasympathetic drive during exercise with increasing levels of physical fitness displayed in the SPPB, which was denoted by the inverse correlation of the later with PNS index during (r = -0.375, p = 0.001) and after exercise (r = -0.229, p = 0.049), in addition to RMSSD during exercise (r = -0.294, p = 0.010)
+
+```{r}
+#| echo: false
+#| warning: false
+#| error: false
+#| fig-width: 8
+#| fig-height: 8
+#| fig-dpi: 300
+dat_long <- melt.data.table(
+  data = dat,
+  id.vars = c("id_record", "tm_pasos", "sppb_total"), 
+  measure.vars = c("hrv_sns_peri", "hrv_stress_peri", "hrv_rmssd_peri", "hrv_sdnn_peri", "hrv_pns_peri")
+)
+
+dat_long[, variable := fcase(
+  grepl("sns", variable), "SNS index",
+  grepl("pns", variable), "PNS index",
+  grepl("stress", variable), "Stress index",
+  grepl("sdnn", variable), "SDNN",
+  grepl("rmssd", variable), "RMSSD"
+)]
+
+dat_long[, variable := factor(variable, levels = c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index"))]
+
+fig3a <- ggplot(dat_long, aes(value, tm_pasos)) +
+  facet_grid(~ variable, scales = "free_x") +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  geom_text(mapping = aes(label = text), data = data.frame(
+    variable = as.factor(c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index")),
+    tm_pasos = c(125, 125, 125, 125, 125),
+    value = c(13, 20, -2, 8, 45),
+    text = c("p = 0.014", "p = 0.141", "p = 0.002", "p = 0.022", "p = 0.261")
+  ), size = 3.5, fontface = "italic") +
+  labs(y = "TMST steps", x = NULL) +
+  scale_y_continuous(expand = c(.1,.1)) +
+  tidybayes::theme_ggdist() +
+  theme(axis.line.x = element_blank(), axis.ticks.x = element_blank(), 
+        axis.text.x = element_blank())
+
+fig3b <- ggplot(dat_long, aes(value, sppb_total)) +
+  facet_grid(~ variable, scales = "free_x") +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  geom_text(mapping = aes(label = text), data = data.frame(
+    variable = as.factor(c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index")),
+    sppb_total = c(14, 14, 14, 14, 14),
+    value = c(13, 20, -2, 8, 45),
+    text = c("p = 0.010", "p = 0.419", "p < 0.001", "p = 0.006", "p = 0.095")
+  ), size = 3.5, fontface = "italic") +
+  labs(y = "SPPB score", x = "Values") +
+  scale_y_continuous(expand = c(.1,.1)) +
+  tidybayes::theme_ggdist()
+
+ggpubr::ggarrange(
+  fig3a, fig3b, ncol = 1
+)
+```
+**Figure 3**. Bivariate dispersion plots between physical fitness and HRV related measures. Significance values for the Pearson's product-moment correlation test is shown.
 
@@
 
-Fig. X. Diagram of main results.
-
-****Insertar Figura resumen resultados****
-
-Footnote: The figure represents the main results of this study.
+![](figures/fig_esquema.jpg)
+**Figure 4**.
+Diagram of main results observed form the data.
 Blue arrows and boxes represent the correlation between variables, while orange arrows and boxes represent the factors mediating the above correlations.

### 2024-03-09 docs/manuscript/manuscript.qmd
minor fix
minor change in citation

@@
 
-Main results are represented in the Figure X.
-
 ![](figures/fig_esquema.jpg)
@@
 We posit that there exists further factors influencing the mediation effect.
-For instance, body distribution of the adipose tissue was not explored, specially when it is known that visceral fat has an important impact on cardiovascular health and its autonomic regulation [@pmid:31553779; pmid:32158768].
+For instance, body distribution of the adipose tissue was not explored, specially when it is known that visceral fat has an important impact on cardiovascular health and its autonomic regulation [@pmid:31553779; @pmid:32158768].
 Participants with higher fitness may have a different fat distribution compared with lower performance individuals, which may induce different effects on autonomic drive to the heart [@pmid:19752873].

### 2024-03-11 docs/manuscript/manuscript.qmd
a citation was corrected

@@
 However, the mediating analysis showed that total muscle mass diminished the main observed effect of body fat on physical fitness, with more muscled individuals displaying a better physical performance despite the body fat percentage.
-This finding suggests that older adults could be benefited of activities that promotes their muscle mass conservation, even when they have high body fat percentages [pmid:27663845].
+This finding suggests that older adults could be benefited of activities that promotes their muscle mass conservation, even when they have high body fat percentages [@pmid:27663845].
 Nevertheless, caution is advised in the interpretation of this result, as the close physiological association among increasing body fat and decreasing muscle mass during aging have been well-studied [@pmid:33335909]

### 2024-06-30 docs/manuscript/manuscript.qmd
update on figures (r-values) and data (age)

@@
 81 community-dwelling older adults were recruited and selected by non-probabilistic sampling, from the Region of Magallanes and Chilean Antartica, Chile.
-They were included if: 65 years age or older; residents of the Region of Magallanes and Chilean Antarctica, Chile; and being able to understand the study aims and instructions.
+They were included if: 55 years age or older; residents of the Region of Magallanes and Chilean Antarctica, Chile; and being able to understand the study aims and instructions.
 But they were excluded in case of: diagnosis of congenital heart disease; consumption of beta-blocker drugs, taking stimulant substances within 24 h before assessments session; motor or cognitive disability; inability to understand instructions or written content; or presence of pain during cardiac or physical assessments.
@@
   data = dat,
-  id.vars = c("id_record", "cc_grasa_total_porcentaje", "cc_masa_muscular_total"), 
+  id.vars = c("id_record", "cc_grasa_total_porcentaje", "cc_masa_muscular_total", "sd_sexo"), 
   measure.vars = c("tm_pasos", "sppb_total")
@@
   facet_grid(~ variable, scales = "free_x") +
-  geom_point() +
+  geom_point(aes(col = sd_sexo)) +
   geom_smooth(method = "lm") +
@@
     value = c(42, 4),
-    text = c("p = 0.013", "p = 0.015")
+    text = c("p = 0.013\nr = -0.279", "p = 0.015\nr = -0.273")
   ), size = 3.5, fontface = "italic") +
@@
     value = c(42, 4),
-    text = c("p = 0.541", "p = 0.584")
+    text = c("p = 0.541\nr = -0.070", "p = 0.584\nr = -0.063")
   ), size = 3.5, fontface = "italic") +
@@
   data = dat,
-  id.vars = c("id_record", "tm_pasos", "sppb_total"), 
+  id.vars = c("id_record", "tm_pasos", "sppb_total", "sd_sexo"), 
   measure.vars = c("hrv_sns_peri", "hrv_stress_peri", "hrv_rmssd_peri", "hrv_sdnn_peri", "hrv_pns_peri")
@@
     value = c(13, 20, -2, 8, 45),
-    text = c("p = 0.014", "p = 0.141", "p = 0.002", "p = 0.022", "p = 0.261")
+    text = c("p = 0.014\nr = -0.285", "p = 0.141\nr = -0.173", "p = 0.002\nr = -0.344", "p = 0.022\nr = 0.265", "p = 0.261\nr = 0.132")
   ), size = 3.5, fontface = "italic") +
@@
     value = c(13, 20, -2, 8, 45),
-    text = c("p = 0.010", "p = 0.419", "p < 0.001", "p = 0.006", "p = 0.095")
+    text = c("p = 0.010\nr = -0.294", "p = 0.419\nr = -0.095", "p < 0.001\nr = -0.375", "p = 0.006\nr = 0.313", "p = 0.095\nr = 0.194")
   ), size = 3.5, fontface = "italic") +

### 2024-06-30 data-raw/data-raw-process.R

@@ -286,6 +286,10 @@ mindfulness[id_record == 36 & id_event_name == "2-meses", hrv_vlf_peri := NA]
 ### No extreme outliers with z-scores greater than 6 beyond this point
 rm(ind, cols, rows, z_score, std_vars, outliers)
 
+
+### Fixing age typo-error
+mindfulness[sd_edad < 60, sd_edad := sd_edad + 10]
+
 # Data export -------------------------------------------------------------
 
 save(mindfulness, file = "data/mindfulness.RData")

### 2024-06-30 docs/manuscript/figures/fig-1.R

@@ -30,7 +30,7 @@ fig <- ggplot(dat_long, aes(value, cc_grasa_total_porcentaje)) +
     variable = as.factor(c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index")),
     cc_grasa_total_porcentaje = c(61, 61, 61, 61, 61),
     value = c(13, 20, -2, 8, 45),
-    text = c("p = 0.029", "p = 0.020", "p = 0.262", "p = 0.030", "p = 0.026")
+    text = c("p = 0.029\nr = 0.253", "p = 0.020\nr = 0.269", "p = 0.262\nr = 0.132", "p = 0.030\nr = -0.252", "p = 0.026\nr = -0.258")
   ), size = 3.5, fontface = "italic") +
   labs(y = "Body fat (%)", x = "Values") +
   tidybayes::theme_ggdist()

### 2024-06-30 docs/manuscript/figures/fig-2.R

@@ -27,7 +27,7 @@ fig2a <- ggplot(dat_long, aes(value, cc_grasa_total_porcentaje)) +
     variable = as.factor(c("TMST steps", "SPPB total score")),
     cc_grasa_total_porcentaje = c(60, 60),
     value = c(42, 4),
-    text = c("p = 0.013", "p = 0.015")
+    text = c("p = 0.013\nr = -0.279", "p = 0.015\nr = -0.273")
   ), size = 3.5, fontface = "italic") +
   labs(y = "Body fat (%)", x = NULL) +
   scale_y_continuous(expand = c(.1,.1)) +
@@ -41,7 +41,7 @@ fig2b <- ggplot(dat_long, aes(value, cc_masa_muscular_total)) +
     variable = as.factor(c("TMST steps", "SPPB total score")),
     cc_masa_muscular_total = c(65.5, 65.5),
     value = c(42, 4),
-    text = c("p = 0.541", "p = 0.584")
+    text = c("p = 0.541\nr = -0.070", "p = 0.584\nr = -0.063")
   ), size = 3.5, fontface = "italic") +
   labs(y = "Muscle mass (kg)", x = NULL) +
   scale_y_continuous(expand = c(.1,.1)) +

### 2024-06-30 docs/manuscript/figures/fig-3.R

@@ -8,7 +8,7 @@ dat <- mindfulness[id_event_name == "Basal", -2L]
 
 dat_long <- melt.data.table(
   data = dat,
-  id.vars = c("id_record", "tm_pasos", "sppb_total"), 
+  id.vars = c("id_record", "tm_pasos", "sppb_total"),
   measure.vars = c("hrv_sns_peri", "hrv_stress_peri", "hrv_rmssd_peri", "hrv_sdnn_peri", "hrv_pns_peri")
 )
 
@@ -30,12 +30,12 @@ fig3a <- ggplot(dat_long, aes(value, tm_pasos)) +
     variable = as.factor(c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index")),
     tm_pasos = c(125, 125, 125, 125, 125),
     value = c(13, 20, -2, 8, 45),
-    text = c("p = 0.014", "p = 0.141", "p = 0.002", "p = 0.022", "p = 0.261")
+    text = c("p = 0.014\nr = -0.285", "p = 0.141\nr = -0.173", "p = 0.002\nr = -0.344", "p = 0.022\nr = 0.265", "p = 0.261\nr = 0.132")
   ), size = 3.5, fontface = "italic") +
   labs(y = "TMST steps", x = NULL) +
   scale_y_continuous(expand = c(.1,.1)) +
   tidybayes::theme_ggdist() +
-  theme(axis.line.x = element_blank(), axis.ticks.x = element_blank(), 
+  theme(axis.line.x = element_blank(), axis.ticks.x = element_blank(),
         axis.text.x = element_blank())
 
 fig3b <- ggplot(dat_long, aes(value, sppb_total)) +
@@ -46,7 +46,7 @@ fig3b <- ggplot(dat_long, aes(value, sppb_total)) +
     variable = as.factor(c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index")),
     sppb_total = c(14, 14, 14, 14, 14),
     value = c(13, 20, -2, 8, 45),
-    text = c("p = 0.010", "p = 0.419", "p < 0.001", "p = 0.006", "p = 0.095")
+    text = c("p = 0.010\nr = -0.294", "p = 0.419\nr = -0.095", "p < 0.001\nr = -0.375", "p = 0.006\nr = 0.313", "p = 0.095\nr = 0.194")
   ), size = 3.5, fontface = "italic") +
[…]

### 2024-08-05 docs/manuscript/manuscript.qmd
paper review update, still under review

@@
 
-**Authors**: Autor,^1^ Autor,^2^ Autor,^3^ Autor,^4^ Autor,^5^ Autor.^6^ \*
+**Authors**: Diego Mabe-Castro^1,2^, Matías Castillo-Aguilar^1^, Matías Mabe-Castro^1,3^, Ruby Méndez Muñoz^1^, Carla Basualto-Alarcón^4,5^, Cristian Núñez-Espinosa^1,3,6^.
 
 ::: {custom-style="Afiliation"}
-^1^ Affiliation, location.
+^1^ Centro Asistencial Docente e Investigación, Universidad de Magallanes, Punta Arenas, Chile
 
-^2^ Affiliation, location.
+^2^ Departamento de Kinesiología, Universidad de Magallanes, Punta Arenas, Chile
 
-^3^ Affiliation, location.
+^3^ Escuela de Medicina. Universidad de Magallanes, Punta Arenas, Chile
 
-^4^ Affiliation, location.
+^4^ Laboratory of Cellular Physiology and Metabolism; Centro PUENTES de Investigación en Salud; Health Sciences Department, University of Aysén, Coyhaique, Chile
 
-^5^ Affiliation, location.
+^5^ Anatomy and Legal Medicine Department, Faculty of Medicine, University of Chile, Santiago, Chile.
 
-^6^ Affiliation, location.
+^6^ Interuniversity Center for Healthy Aging
 :::
 
-### \*Corresponding author
+::: {custom-style="Afiliation"}
+Corresponding Author:
 
-Autor, Affiliation, location. e-mail: autor\@mail.com. 
-Address: XXXX. Phone: XXXX.
+Cristian Núñez-Espinosa
+
+Avenida Bulnes 01855, Box 113-D, Punta Arenas, Región de Magallanes y la Antártica Chilena, Chile
+
+Email address: Cristian.nunez\@umag.cl
+:::
 
@@
 
-**Background**: Physical fitness, body composition, and the autonomic response to exercise are important predictors of health-related outcomes in older adults, but their interplay is not fully understood, therefore, this study aims to assess the associations between these variables and possible mediating effects.
-**Methods**: an observational cross-sectional study was conducted in 81 community-dwelling older people, recruited by non-probabilistic sampling.
-To obtain body composition parameters, a bio-impedance analysis was made, then, the Short Physical Performance Battery (SPPB) was used to assess physical fitness, and the Heart Rate Variability (HRV) was recorded before, during and after the Two-Minute Step Test (TMST), to assess cardiac autonomic response to physical exercise.
-Geriatric depression and anxiety symptoms were assessed as secondary measures.
-Correlation and mediating effects analyses were carried out.
-**Results**: SPPB score was negatively related to body fat percentage (r = -0.273, p = 0.015), but positively related to SNS index during exercise (r = 0.313, p = 0.006).
-Also, a negative link existed between body fat and sympathetic activity: SNS index (r = -0.252, p = 0.030) and Stress index (r = 0.258, p = 0.027) during TMST.
-But a positive link with parasympathetic activity: RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020) **Conclusion**: Body fat was inversely correlated to both physical fitness and cardiac autonomic activity during exercise, with sympathetic drive associated with physical fitness in older people.
-Furthermore, physical fitness may mediate the effect of body fat on cardiac autonomic activity during exercise, and total muscle mass may mitigate the negative effect of increased adiposity on physical fitness, highlighting its pivotal role in older people's health.
-However, further research is needed to elucidate mediating effects of psychological variables in this field.
+**Background**.
+Age-related changes in body composition affect physical fitness in older adults.
+However, whether the autonomic response is associated with body fat percentage and its implication for physical fitness is not fully understood.
 
-**Keywords**: Aged, Physical Functional Performance, Exercise Test, Body Composition, Autonomic Nervous System
+**Aim**. To understand the association between physical fitness, body composition, and heart rate variability in older people and its mediating factors.
+
+**Methods**.
+A cross-sectional study with 81 older adults was conducted, assessing Short Physical Performance Battery (SPPB), Two-minute Step Test (TMST), body composition, and cardiac autonomic response.
+Correlation and mediation analyses were performed.
+
+**Results**.
+Body fat percentage negatively correlated with physical fitness (SPPB: r = -0.273, p = 0.015; TMST: r = -0.279, p = 0.013) and sympathetic activity (SNS index: r = -0.252, p = 0.030), yet positively correlated with parasympathetic tone (RMSSD: r = 0.253, p = 0.029; SDNN: r = 0.269, p = 0.020).
+Physical fitness associated with sympathetic nervous system index (SPPB: r = 0.313, p = 0.006; TMST: r = 0.265, p = 0.022) and parasympathetic nervous system index (TMST: r = -0.344, p = 0.003).
+Muscle mass mediated body fat's impact on physical fitness, while physical fitness mediated body fat's impact on autonomic response.
+
+**Conclusion**.
+Body composition and cardiac autonomic response to exercise are associated with physical fitness in older people, highlighting a possible protective effect of muscle mass against the decline in physical fitness associated with increased body fat.
 
@@
 
-The process of aging is accompanied by a multitude of physiological changes that have the potential to significantly impact people's overall health and well-being [@pmid:30185958; @pmid:31154942].
-Body composition and cardiovascular function are particularly relevant, as they are closely associated with the aging process [@pmid:36948782; @pmid:36443946; @pmid:30065258; @pmid:36918543; @pmid:32483625].
-In recent times, there has been a growing interest in comprehending the interplay between physical fitness, body composition, and cardiac autonomic response to exercise in older individuals, as these factors play pivotal roles in determining health outcomes within this age group.
+Aging is accompanied by a multitude of physiological changes that have the potential to significantly impact people's overall health and well-being [@pmid:30185958; @pmid:31154942].
+Body composition and cardiovascular function are particularly relevant, as they are closely associated with aging [@pmid:36948782; @pmid:36443946; @pmid:30065258; @pmid:36918543; @pmid:32483625].
+Recently, there has been a growing interest in comprehending the interplay between physical fitness, body composition, and cardiac autonomic response to exercise in older individuals, as these factors play pivotal roles in determining health outcomes within this age group.
 
@@
 
-Physical fitness has been classically defined as the ability to perform daily tasks with vigor and without undue fatigue, and with ample energy to enjoy leisure-time pursuits and to meet unforeseen emergencies [@pmid:3920716].
+Physical fitness is the ability to perform daily tasks with vigor, without undue fatigue, and with ample energy to enjoy leisure-time pursuits and meet unforeseen emergencies [@pmid:3920716].
 Therefore, it is a crucial component of healthy aging.
-The term encompasses a multitude of aspects, typically including cardiorespiratory endurance, muscular endurance and strength, body composition, and flexibility [@pmid:3920711].
-Numerous studies have demonstrated the positive impact of physical fitness on overall health and longevity, emphasizing its role in reducing the risk of chronic diseases, and improving quality of life and functional independence in older adults [@pmid:9177581; @pmid:29293447; @pmid:30768197; @pmid:36332759]
+The term encompasses many aspects, including cardiorespiratory endurance, muscular endurance and strength, body composition, and flexibility [@pmid:3920711].
+Numerous studies have demonstrated the positive impact of physical fitness on overall health and longevity, emphasizing its role in reducing the risk of chronic diseases and improving quality of life and functional independence in older adults [@pmid:9177581; @pmid:29293447; @pmid:30768197; @pmid:36332759]
 
@@
 
-Autonomic nervous system (ANS) plays a critical role in regulating cardiovascular function, with sympathetic and parasympathetic branches exerting opposing effects on heart rate and vascular tone.
+The autonomic nervous system (ANS) plays a critical role in regulating cardiovascular function, with sympathetic and parasympathetic branches exerting opposing effects on heart rate and vascular tone.
 Furthermore, during stressful situations, such as physical exercise, the ANS ensures a sufficient cardiac response to higher metabolic demands [@pmid:16627049].
-The measurement of heart rate variability (HRV), serves as a non-invasive indicator of ANS activity and cardiovascular health [@pmid:37641564; @pmid:29033120].
-A reduction in rest HRV has been linked to various adverse outcomes, including cardiovascular events and morbidity, while insufficient HRV reduction during exercise may lead to impaired physical capacity [@pmid:29033120; @pmid:33390146; @pmid:36054204].
+Heart rate variability (HRV) is a non-invasive indicator of ANS activity and cardiovascular health [@pmid:37641564; @pmid:29033120].
+A reduction in resting HRV has been linked to various adverse outcomes, including cardiovascular events and morbidity, while insufficient HRV reduction during exercise may lead to impaired physical capacity [@pmid:29033120; @pmid:33390146; @pmid:36054204].
 
@@
 Psychological factors such as anxiety and depression have also been linked to changes in ANS activity [@doi:10.1016/j.jad.2018.04.071; @pmid:35340102].
-Notably, older individuals may be more susceptible to these psychological factors, making it crucial to measure and control for their effects [@pmid:38007499; @pmid:25980510].
+Older individuals may be more susceptible to these psychological factors, making it crucial to measure and control for their effects [@pmid:38007499; @pmid:25980510].
 
@@
 
-Despite extensive research into physical fitness, body composition, and cardiovascular health in older adults, gaps persist in the understanding of their interrelationships and factors mediating them.
+Despite extensive research into physical fitness, body composition, and cardiovascular health in older adults, gaps persist in understanding their interrelationships and the factors mediating them.
 One area of interest is the relationship between physical fitness, body fat percentage, and cardiac autonomic response to exercise in older individuals.
@@
 
-The conceptualization of this investigation is based on the recognition of the interplay between physical fitness, body composition, and autonomic cardiovascular regulation in aging.
-We aim to answer the following research question: How does physical fitness relate to body fat percentage and cardiac autonomic response to exercise in older people?
+The conceptualization of this investigation is based on recognizing the interplay between physical fitness, body composition, and autonomic cardiovascular regulation in aging.
+We aim to answer the following research question: How does physical fitness relate to body fat percentage and cardiac autonomic responses to exercise in older people?
 By elucidating these correlations, we hope to gain valuable insights into the relationships that underlie age-related modifications in physical fitness and identify potential interventions to enhance health outcomes in older adults.
-Even more, we seek to understand the mediating effect of the variables, including frequent psychological traits, such as geriatric depression and anxiety.
+Even more, we seek to understand the mediating effect of the variables, including physical aspects and frequent psychological traits, such as geriatric depression and anxiety.
 
@@
 
-We hypothesize that physical fitness will be inversely related to body fat percentage and associated with cardiac autonomic response during exercise in older individuals.
-In our opinion, it will be essential for this population to have an optimal cardiac autonomic response, characterized by a shift towards sympathetic activity predominance, in order to achieve a grater performance during physical fitness testing.
-
-The primary objective of this study is to investigate the relationship between physical fitness, body fat percentage, and cardiac autonomic response to exercise in older people.
-Even more, We aimed to assess the mediating effect of psychological variables on these relationships, with implications for cardiovascular health and functional capacity in aging.
+We hypothesized that physical fitness would be inversely related to body fat percentage and associated with cardiac autonomic response during exercise in older individuals.
+We believe an optimal cardiac autonomic response, characterized by a shift towards sympathetic activity predominance, will be essential for this population to achieve greater performance during physical fitness testing.
 
@@
 
-A prospective observational, cross-sectional study was conducted  in one data collection session, to determine the associations between physical fitness, body composition and cardiac autonomic response to moderate-intensity exercise.
+A prospective observational, cross-sectional study was conducted in one data collection session to determine the associations between physical fitness, body composition, and cardiac autonomic response to moderate-intensity exercise.
 
@@
 
-This study was conducted at the Centro Asistencial Docente e Investigación (Assistance, Teaching, and Research Center), belonging to the University of Magallanes (CADI-UMAG), in the city of Punta Arenas, Chile.
-All the assessments were made between 09:00 and 11:00 hours of the morning.
-The privacy and comfort of the subjects was ensured, the room temperature was set at 20°C and a white artificial lighting was used.
+This study was conducted at the Centro Asistencial Docente e Investigación (Assistance, Teaching, and Research Center), which belongs to the University of Magallanes (CADI-UMAG) in Punta Arenas, Chile.
+All the assessments were made between 09:00 and 11:00 a.m.
+The privacy and comfort of the subjects were ensured; the room temperature was set at 20°C, and white artificial lighting was used.
 
@@
 
-81 community-dwelling older adults were recruited and selected by non-probabilistic sampling, from the Region of Magallanes and Chilean Antartica, Chile.
-They were included if: 55 years age or older; residents of the Region of Magallanes and Chilean Antarctica, Chile; and being able to understand the study aims and instructions.
-But they were excluded in case of: diagnosis of congenital heart disease; consumption of beta-blocker drugs, taking stimulant substances within 24 h before assessments session; motor or cognitive disability; inability to understand instructions or written content; or presence of pain during cardiac or physical assessments.
+81 community-dwelling older adults were recruited and selected by non-probabilistic sampling from the Region of Magallanes and Chilean Antarctica, Chile.
+If they were 60 years or older, residents of the Region of Magallanes and Chilean Antarctica, Chile, they were included and understood the study aims and instructions.
+However, they were excluded in the case of diagnosis of congenital heart disease, consumption of beta-blocker drugs, taking stimulant substances within 24 h before assessment session, motor or cognitive disability, inability to understand instructions or written content, or presence of pain during cardiac or physical assessments.
 
-All participating subjects gave their permission and provided informed consent before participation.
-The Ethics Committee of the University of Magallanes (N°XXX/YY/YEAR) approved this study following the regulations established by the Declaration of Helsinki on ethical principles in human beings.
+All participants gave their permission and provided informed written consent before participation.
+The Ethics Committee of the University of Magallanes (N°10/CEC-UMAG/2023) approved this study, following the regulations established by the Declaration of Helsinki on ethical principles in human beings.
+The flow diagram of the included participants can be seen in Figure 1.
 
@@
 
-During recruitment, the participants were instructed to avoid the use of psychoactive substances for the 24 hours prior to the assessment, and to sleep for at least 7 hours during the previous night.
-Upon arrival, participants were informed about the study aims and risks associated with their participation, and the socio-demographic and medical information was collected during the initial interview.
+During recruitment, the participants were instructed to avoid the use of psychoactive substances for 24 hours before the assessment and to sleep for at least 7 hours the previous night.
+Upon arrival, participants were informed about the study aims and risks associated with their participation, and sociodemographic and medical information was collected during the initial interview.
 Then, Body composition parameters were measured using bioimpedance analysis, and psychological questionnaires for anxiety and depression screening were applied by a supervised psychology undergraduate student.
@@
 
-HRV was recorded through a non-invasive chest band, immediately before, during, and immediately after the execution of the two-minutes step test (TMST).
-Vital signs, including blood pressure, were monitored throughout the test, and participants' well-being was visually checked to ensure they were comfortable and prepared.
-For resting HRV measurements (before and after the application of the TMST), the volunteers remained seated in a chair, with feet and back supported, ensuring avoiding to talk during the recordings.
-R-R intervals were recorded continuously during the last 10 min of rest, and were subsequently analyzed for 5 min in each occasion.
+HRV was recorded through a non-invasive chest band immediately before, during, and after executing the two-minute step test (TMST).
+Vital signs, including blood pressure, were monitored throughout the test, and participants’ well-being was visually checked to ensure they were comfortable and prepared.
+For resting HRV measurements (before and after the application of the TMST), the volunteers remained seated in a chair, with feet and back supported, ensuring avoidance of talking during the recordings.
+R-R intervals were recorded continuously during the last 10 min of rest and were analyzed for 5 min on each occasion.
 The breathing rate was spontaneous.
-As part of the protocol, it was ensured that the participant needed to have a blood pressure less than 140/90 mmHg to start the HRV measurements.
+As part of the protocol, it was ensured that the participant had a blood pressure of less than 140/90 mmHg to start the HRV measurements.
 
-Finally, after 15 minutes of resting from the TMST, the Short Physical Performance Battery (SPPB) was applied to the participants.
-Blood pressure, heart rate and general aspect of the subjects was verified before ending the session.
+Finally, after 15 minutes of resting from the TMST, the Short Physical Performance Battery (SPPB) was administered to the participants.
+Before the session ended, the subjects ' blood pressure, heart rate, and general appearance were verified.
 
@@
 
-Balance: to achieve the maximal score (4 points), the subject should be able to stay balanced for at least 10 seconds in side-by-side stand (feet together), semi-tandem stand and tandem stand.
+Balance: to achieve the maximal score (4 points), the subject should be able to stay balanced for at least 10 seconds in a side-by-side stand (feet together), semi-tandem stand, and tandem stand.
 
-Usual gait speed: the subjects are asked to walk normally a 4 meters distance.
-Two chances are given and the best of both is registered.
+Usual gait speed: The subjects are asked to walk normally at a 4-meter distance.
+They are given two chances, and the best of both is registered.
 To achieve the maximal score (4 points), they should obtain 4.82 sec.
@@
 
-Lower body fitness: The subjects are asked to perform 5 chair stands without using their arms and time upon complete is registered. A time lower or equal 11.19 sec. gives the maximal score (4 points).
+Lower body fitness: The subjects are asked to perform five chair stands without using their arms, and time is registered upon completion.
+A time lower or equal to 11.19 sec.
+gives the maximal score (4 points).
 
-To obtain the final score, the summing of the three components is calculated, with a maximal total score of 12 points.
+To obtain the final score, the sum of the three components is calculated, with a maximal total score of 12 points.
 
@@
 
-Cardiac autonomic activity was assessed using R-R interval recordings (HRV) obtained through the Polar Team2 system (Polar®) application.
-Artifacts and ectopic heartbeats, which did not exceed 3% of the recorded data, were excluded [@malik1996heart].
+The cardiac autonomic activity was assessed using R-R interval recordings (HRV) obtained through the Polar Team2 system (Polar®) application.
+Artifacts and ectopic heartbeats were excluded, not exceeding 3% of the recorded data [@malik1996heart].
 Time-domain parameters considered for analysis included the square root of the mean squared differences of successive R-R intervals (RMSSD, expressed in ms) as an index of parasympathetic activity [@pmid:20033207], and the standard deviation of RR intervals (SDNN), reflecting total variability encompassing both sympathetic and parasympathetic contributions to cardiac autonomic function [@pmid:9401419; @buchheit2006cardiac].
-Additionally, the Stress Index (SI) and Parasympathetic and Sympathetic Nervous System Index (PNS and SNS, respectively) were computed.
+The Stress Index (SI) and Parasympathetic and Sympathetic Nervous System Index (PNS and SNS) were computed.
 The PNS Index, indicative of total vagal stimulation, was derived from mean R-R intervals, RMSSD, and Poincaré Plot Index SD1 in normalized units (linked to RMSSD), representing deviations from normal population averages [@pmid:9401419; @GSID:XkAQwTGLgNsJ].
 The SNS Index, reflecting total sympathetic stimulation, was derived from mean R-R intervals, Baevsky’s Stress Index (positively related to cardiovascular system stress and cardiac sympathetic activity), and the Poincaré Plot Index SD2 in normalized units (related to SDNN) with interpretation similar to the PNS Index [@pmid:9401419; @GSID:XkAQwTGLgNsJ].
-The SI serves as an indicator representing the ANS control system's workload [@doi:10.1159/000513781], normalized by the square root of Baevsky’s SI [@2009MethodicalRU].
-All analyses conducted to compute HRV related indices were performed using Kubios HRV® software (Kuopio, Finland).
+The SI indicates the ANS control system’s workload [@doi:10.1159/000513781], normalized by the square root of Baevsky’s SI [@2009MethodicalRU].
+All analyses were conducted to compute HRV-related indices using Kubios HRV® software (Kuopio, Finland).
 
@@
 The TMST is a sub-test from the Senior Fitness Test, demanding a low to moderate intensity [@Rikli1999; @ijerph20054456].
-The test consists of a two-minute assessment designed to evaluate cardiorespiratory fitness.
-Participants are instructed to raise their right knee to a marked point on a wall as many times as possible within the given time frame, ensuring that each raise reaches at least a 70° angle at the thigh-femoral joint.
-The number of valid steps were recorded for each subject.
+It consists of a two-minute assessment designed to evaluate cardiorespiratory fitness.
+Participants are instructed to raise their right knee to a marked point on a wall as often as possible within the given time frame, ensuring that each raise reaches at least a 70° angle at the thigh-femoral joint.
+The number of valid steps was recorded for each subject.
 
@@
 
-The 30-questions Geriatric Depression Scale (GDS-30) was employed to assess subject's depressive symptoms [@pmid:27633329].
-It consists of a 30 dichotomous questionnaire, where participants are asked about their past-week feelings in reference to depressive symptoms (for instance, "Do you feel that your life is empty?"), with higher scores (i.
-e., "yes" responses) representing more depressive symptoms.
+The 30-question Geriatric Depression Scale (GDS-30) was employed to assess the subject’s depressive symptoms [@pmid:27633329].
+It consists of a dichotomous questionnaire, where participants are asked about their past-week feelings about depressive symptoms (for instance, “Do you feel that your life is empty?”), with higher scores (i.e., “yes” responses) representing more depressive symptoms.
 It was first developed by Yesavage and colleagues [@pmid:7183759] and is currently widely used [@pmid:34922595].
-The spanish version, employed in this study, has been previously validated, with a Cronbach alpha coefficient of 0.82 [@pmid:11921157].
+The Spanish version employed in this study has been previously validated, with a Cronbach alpha coefficient of 0.82 [@pmid:11921157].
+The authors have permission to use this instrument from the copyright holders.
 
@@
 
-The Beck anxiety Inventory was employed to assess anxiety symptoms.
-It consists of a 21-items questionnaire based on usual anxiety symptoms and a 4-options Likert scale from 0 ("Not at all") to 3 ("Severely"), meaning the severity reported by the subject in each one.
+The Beck Anxiety Inventory was employed to assess anxiety symptoms.
+It consists of a 21-item questionnaire based on usual anxiety symptoms and a 4-option Likert scale from 0 (“Not at all”) to 3 (“Severely”), meaning the severity reported by the subject in each one.
 It was originally developed by Beck and colleagues [@pmid:3204199].
-In this study, the spanish version is used, which has demonstrated a high internal consistency in older people (α = 0.94) [@RODRIGUEZREYNALDO2001353].
+This study uses the Spanish version, demonstrating a high internal consistency in older people (α = 0.94) [@RODRIGUEZREYNALDO2001353].
+The authors have permission to use this instrument from the copyright holders.
 
@@
 
-We used mean and standard deviation (SD) to describe continuous variables, absolute and relative frequencies to describe discrete variables.
-In order to assess the relationship between variables, we used Pearson's product-moment correlation (r), and to assess differences between groups we used standardized mean difference (SMD) and 95% confidence intervals (CI~95%~).
+We used mean and standard deviation (SD) to describe continuous variables and absolute and relative frequencies to describe discrete variables.
+We used Pearson’s product-moment correlation (r) to assess the relationship between variables.
+To assess differences between groups, we used standardized mean difference (SMD) and 95% confidence intervals (CI!95%!).
 
 As a way of controlling for the influence of psycho-physiological variables on the cardiac autonomic response of the subjects to exercise, we assess the average causal mediation effect (ACME), the average direct effect (ADE) of the main effects after taking into account the effect of moderator variables into the observed relationships and the proportion of the effect that its mediated by these variables [@kosuke2010identification].
-To assess the significance of moderating variables, we used non overlapping CI~95%~, estimated based on non parametric bootstrapping by using Monte Carlo resampling, and bias-corrected and accelerated CI~95%~, using the *mediation* R package to this end [@tingley2014mediation].
+TTo assess the significance of moderating variables, we used nonoverlapping CI95%, estimated based on nonparametric bootstrapping using Monte Carlo resampling, and bias-corrected.
+We accelerated CI95%, using the *mediation* R package to this end [@tingley2014mediation].
 
-We defined a type I error rate at 5% ($\alpha$ = 0.05) as our threshold for null hypothesis significance testing, and non-overlapping CI~95%~ over the null effect for sex differences and mediation analyses.
-All analyses were computed using the *R* programming language for statistical computing, [@rlanguage].
+We defined a type I error rate of 5% (α = 0.05) as our threshold for null hypothesis significance testing and nonoverlapping CI95% over the null effect for sex differences and mediation analyses.
+To estimate the confidence in our conclusions and, consequently, the statistical power of our analyses, we calculated the probability of correctly rejecting the null hypothesis for the correlation tests post hoc, considering a low (i.e., r = 0.3) to moderate effect size (r = 0.5)
+
+All analyses were computed using the R programming language for statistical computing [@rlanguage].
 
@@
 
-The sample assessed consisted on 81 adults (age: 71.1 ± 6.2 years old, BMI: 31 ± 6.2 kg/m^2^), of whom 67 (82.7%) were females.
-Sociodemographic and anthropometric characteristics of the sample can be seen in table 1.
+The sample consisted of 81 adults (71.1 ± 6.2 years old, BMI: 31 ± 6.2 kg/m^2^) and 67 (82.7%) females. Table 1 shows the characteristics of the sample.
 
@@
 ```{r}
+#| eval: false
 #| echo: false
@@
 ```
-**Table 1**.
-Main sample characteristics.
-Body composition and sociodemograpic variables are displayed for the overall sample and aggregated by sex.
-Differences between males and females are displayed as SMD and CI~95%~ for continuous variables.
 
@@
 We observed that body fat percentage was linked to a lower SNS (r = -0.252, p = 0.030) and Stress index (r = -0.258, p = 0.027) response during the TSMT.
-In addition, body fat percentage was positively correlated with time domain HRV during exercise, more specifically RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020).
-No other body composition variables were found to influence the cardiac autonomic response to exercise.
+In addition, body fat percentage was positively correlated with time domain HRV during exercise, specifically RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020) (See Figure 2).
+No other body composition variables correlate with the cardiac autonomic response to exercise.
 
@@
   facet_grid(~ variable, scales = "free_x") +
-  geom_point(aes(col = sd_sexo)) +
+  geom_point() +
   geom_smooth(method = "lm") +

### 2024-08-05 to-do.qmd
paper review update, still under review

---
title: "Adressing reviewers comments"
format: pdf
editor_options: 
  chunk_output_type: console
execute: 
  eval: false
---

```{r}
## Load packages
library(data.table)

## Load data
load("data/mindfulness.RData")

## Filtering study data
dat <- mindfulness[id_event_name == "Basal", -2L]

## Add equilibrium category 
dat[, sppb_equil_cat := sppb_equil_bip + sppb_equil_semitandem + sppb_equil_tandem]
```

## To-do

- [x] "Run separate correlations on an only female sample and add to the manuscript to identify exactly how males influence the results".
- [x] "The correlations for each individual SPPB assessment results (i.e., time of 4-meter walk, time of repeated chair rise, etc.) should be reported".

# Body composition and autonomic response

## Result 1: more body fat percentage was linked to a lower SNS response to TSMT

```{r}
## Combined data
dat[, cor.test(cc_grasa_total_porcentaje, hrv_sns_peri) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_grasa_total_porcentaje, hrv_sns_peri) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: r = -0.252, p = 0.030 \*
- Males: r = 0.174, p = 0.589
- Females: r = -0.356, p = 0.005 \*\*

## Result 2: more body fat percentage was linked to a lower SI response to TSMT

```{r}
## Combined data
dat[, cor.test(cc_grasa_total_porcentaje, hrv_stress_peri) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_grasa_total_porcentaje, hrv_stress_peri) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: r = -0.258, p = 0.027 \*
- Males: r = 0.204, p = 0.525
- Females: r = -0.316, p = 0.012 \*

## Result 3: body fat percentage was positively correlated with RMSSD during exercise

```{r}
## Combined data
dat[, cor.test(cc_grasa_total_porcentaje, hrv_rmssd_peri) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_grasa_total_porcentaje, hrv_rmssd_peri) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: r = 0.253, p = 0.029 \*
- Males: r = -0.259, p = 0.416
- Females: r = 0.246, p = 0.054 \.

## Result 4: body fat percentage was positively correlated with SDNN during exercise

```{r}
## Combined data
dat[, cor.test(cc_grasa_total_porcentaje, hrv_sdnn_peri) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_grasa_total_porcentaje, hrv_sdnn_peri) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: r = 0.269, p = 0.020 \*
- Males: r = -0.058, p = 0.859
- Females: r = 0.247, p = 0.053 \.

# Body composition and physical fitness

## Results 1: greater SPPB score witb lower body fat

Note: Using Spearman's rho as the constituents of the SPPB domains are ordinal (levels from 1 to 4) and body fat percentage is continuous.

#### sppb_equil_cat

```{r}
## Combined data
dat[, cor.test(cc_grasa_total_porcentaje, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_grasa_total_porcentaje, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = 0.052, p = 0.656
- Males: rho = 0.044, p = 0.893
- Females: rho = 0.017, p = 0.896

#### sppb_sts_cat

```{r}
## Combined data
dat[, cor.test(cc_grasa_total_porcentaje, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_grasa_total_porcentaje, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.279, p = 0.013 \*
- Males: rho = -0.304, p = 0.337
- Females: rho = -0.347, p = 0.004 \*\*

#### sppb_velocidad_cat

```{r}
## Combined data
dat[, cor.test(cc_grasa_total_porcentaje, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_grasa_total_porcentaje, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.261, p = 0.021 \*
- Males: rho = 0.172, p = 0.593
- Females: rho = -0.309, p = 0.012 \*

## Results 2: greater SPPB score with lower BMI

Note: Using Spearman's rho as the constituents of the SPPB domains are ordinal (levels from 1 to 4) and BMI is continuous.

#### sppb_equil_cat

```{r}
## Combined data
dat[, cor.test(cc_imc, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_imc, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.083, p = 0.474
- Males: rho = 0.1, p = 0.77
- Females: rho = -0.109, p = 0.39

#### sppb_sts_cat

```{r}
## Combined data
dat[, cor.test(cc_imc, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_imc, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.325, p = 0.004 \*\*
- Males: rho = -0.294, p = 0.38
- Females: rho = -0.324, p = 0.008 \*\*

#### sppb_velocidad_cat

```{r}
## Combined data
dat[, cor.test(cc_imc, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_imc, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.305, p = 0.007 \*\*
- Males: rho = -0.108, p = 0.752
- Females: rho = -0.318, p = 0.009 \*\*

## Results 3: greater TMST steps with lower body fat percentage

```{r}
## Combined data
dat[, cor.test(cc_grasa_total_porcentaje, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_grasa_total_porcentaje, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: r = -0.279, p = 0.013
- Males: r = 0.085, p = 0.792
- Females: r = -0.311, p = 0.01

## Results 4: greater TMST steps with lower bodyweight

```{r}
## Combined data
dat[, cor.test(cc_peso_corporal, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_peso_corporal, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: r = -0.232, p = 0.041 \*
- Males: r = -0.198, p = 0.559
- Females: r = -0.275, p = 0.024 \*

## Results 4: greater TMST steps with lower body fat

```{r}
## Combined data
dat[, cor.test(cc_grasa_total_porcentaje, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_grasa_total_porcentaje, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: r = -0.279, p = 0.013 \*
- Males: r = 0.085, p = 0.792
- Females: r = -0.311, p = 0.01 \*\*

## Null result 6: Total SPPB and total muscle mass

Note: Using Spearman's rho as the constituents of the SPPB domains are ordinal (levels from 1 to 4) and total muscle mass is continuous.

#### sppb_equil_cat

```{r}
## Combined data
dat[, cor.test(cc_masa_muscular_total, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_masa_muscular_total, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = 0.016, p = 0.888
- Males: rho = 0.044, p = 0.893
- Females: rho = 0.049, p = 0.699

#### sppb_sts_cat

```{r}
## Combined data
dat[, cor.test(cc_masa_muscular_total, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_masa_muscular_total, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.255, p = 0.024 \*
- Males: rho = -0.126, p = 0.697
- Females: rho = -0.327, p = 0.007 \*\*

#### sppb_velocidad_cat

```{r}
## Combined data
dat[, cor.test(cc_masa_muscular_total, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_masa_muscular_total, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.136, p = 0.236
- Males: rho = 0.215, p = 0.502
- Females: rho = -0.254, p = 0.039 \*

## Null result 7: Total number of steps in TMST and total muscle mass

```{r}
## Combined data
dat[, cor.test(cc_masa_muscular_total, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(cc_masa_muscular_total, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: r = -0.07, p = 0.541
- Males: r = -0.258, p = 0.418
- Females: r = -0.237, p = 0.054 \.

# Physical fitness and cardiac autonomic response to exercise

However, this appears to be accompanied by a proportional decrease in the parasympathetic drive during exercise with increasing levels of physical fitness displayed in the SPPB, which was denoted by the latter's inverse correlation with the PNS index during (r = -0.375, p = 0.001) and after exercise (r = -0.229, p = 0.049), in addition to RMSSD during exercise (r = -0.294, p = 0.010). See Figure 4.

## Results 1: Greater SPPB with greater SNS index peri

Note: Using Spearman's rho as the constituents of the SPPB domains are ordinal (levels from 1 to 4) and SNS index is continuous.

#### sppb_equil_cat

```{r}
## Combined data
dat[, cor.test(hrv_sns_peri, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_sns_peri, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = 0.082, p = 0.492
- Males: rho = 0, p = 1
- Females: rho = 0.121, p = 0.356

#### sppb_sts_cat

```{r}
## Combined data
dat[, cor.test(hrv_sns_peri, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_sns_peri, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = 0.345, p = 0.003 \*\*
- Males: rho = 0.607, p = 0.028 \*
- Females: rho = 0.294, p = 0.021 \*

#### sppb_velocidad_cat

```{r}
## Combined data
dat[, cor.test(hrv_sns_peri, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_sns_peri, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = 0.098, p = 0.407
- Males: rho = -0.137, p = 0.654
- Females: rho = 0.124, p = 0.341

## Results 2: Greater TMST steps with greater SNS index peri

```{r}
## Combined data
dat[, cor.test(hrv_sns_peri, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_sns_peri, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: r = 0.265, p = 0.022 \*
- Males: r = -0.211, p = 0.51
- Females: r = 0.326, p = 0.01 \*

## Results 3: Greater SPPB with greater mean HR peri

Note: Using Spearman's rho as the constituents of the SPPB domains are ordinal (levels from 1 to 4) and mean HR is continuous.

#### sppb_equil_cat

```{r}
## Combined data
dat[, cor.test(hrv_mean_hr_peri, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_mean_hr_peri, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = 0.128, p = 0.28
- Males: rho = 0, p = 1
- Females: rho = 0.143, p = 0.276

#### sppb_sts_cat

```{r}
## Combined data
dat[, cor.test(hrv_mean_hr_peri, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_mean_hr_peri, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = 0.387, p = 0.001 \*\*
- Males: rho = 0.558, p = 0.047 \*
- Females: rho = 0.334, p = 0.009 \*\*

#### sppb_velocidad_cat

```{r}
## Combined data
dat[, cor.test(hrv_mean_hr_peri, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_mean_hr_peri, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = 0.123, p = 0.297
- Males: rho = -0.212, p = 0.486
- Females: rho = 0.213, p = 0.100

## Results 4: Greater TMST steps with greater mean HR peri

```{r}
## Combined data
dat[, cor.test(hrv_mean_hr_peri, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_mean_hr_peri, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: r = 0.338, p = 0.003 \*\*
- Males: r = -0.17, p = 0.598
- Females: r = 0.414, p = 0.001 \*\*

## Results 5: Greater SPPB with less mean R-R interval peri

Note: Using Spearman's rho as the constituents of the SPPB domains are ordinal (levels from 1 to 4) and mean R-R interval is continuous.

#### sppb_equil_cat

```{r}
## Combined data
dat[, cor.test(hrv_mean_rr_peri, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_mean_rr_peri, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.125, p = 0.291
- Males: rho = 0, p = 1
- Females: rho = -0.141, p = 0.284

#### sppb_sts_cat

```{r}
## Combined data
dat[, cor.test(hrv_mean_rr_peri, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_mean_rr_peri, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.394, p = 0.001 \*\*
- Males: rho = -0.558, p = 0.047 \*
- Females: rho = -0.338, p = 0.008 \*\*

#### sppb_velocidad_cat

```{r}
## Combined data
dat[, cor.test(hrv_mean_rr_peri, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_mean_rr_peri, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.125, p = 0.29
- Males: rho = 0.256, p = 0.398
- Females: rho = -0.214, p = 0.098

## Results 6: Greater TMST steps with less mean R-R interval peri

```{r}
## Combined data
dat[, cor.test(hrv_mean_rr_peri, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_mean_rr_peri, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: r = -0.311, p = 0.007 \*\*
- Males: r = 0.16, p = 0.62
- Females: r = -0.398, p = 0.001 \*\*

## Results 7: Greater TMST with less RMSSD peri

```{r}
## Combined data
dat[, cor.test(hrv_rmssd_peri, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_rmssd_peri, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: r = -0.285, p = 0.014 \*
- Males: r = -0.08, p = 0.805
- Females: r = -0.288, p = 0.023 \*

## Results 8: Greater TMST with less PNS index peri

```{r}
## Combined data
dat[, cor.test(hrv_pns_peri, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_pns_peri, tm_pasos) |> 
      with({
        paste0("r = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: r = -0.344, p = 0.003 \*\*
- Males: r = 0.115, p = 0.722
- Females: r = -0.41, p = 0.001 \*\*

## Results 9: Greater SPPB with less PNS index peri

Note: Using Spearman's rho as the constituents of the SPPB domains are ordinal (levels from 1 to 4) and PNS index is continuous.

#### sppb_equil_cat

```{r}
## Combined data
dat[, cor.test(hrv_pns_peri, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_pns_peri, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.109, p = 0.36
- Males: rho = 0.077, p = 0.802
- Females: rho = -0.154, p = 0.24

#### sppb_sts_cat

```{r}
## Combined data
dat[, cor.test(hrv_pns_peri, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_pns_peri, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.407, p < 0.001 \*\*\*
- Males: rho = -0.518, p = 0.07
- Females: rho = -0.363, p = 0.004 \*\*

#### sppb_velocidad_cat

```{r}
## Combined data
dat[, cor.test(hrv_pns_peri, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_pns_peri, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.108, p = 0.362
- Males: rho = 0.216, p = 0.48
- Females: rho = -0.179, p = 0.168

## Results 10: Greater SPPB with less PNS index post index is continuous

Note: Using Spearman's rho as the constituents of the SPPB domains are ordinal (levels from 1 to 4) and PNS index is continuous.

#### sppb_equil_cat

```{r}
## Combined data
dat[, cor.test(hrv_pns_post, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_pns_post, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.226, p = 0.054
- Males: rho = -0.077, p = 0.802
- Females: rho = -0.259, p = 0.046 \*

#### sppb_sts_cat

```{r}
## Combined data
dat[, cor.test(hrv_pns_post, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_pns_post, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.165, p = 0.161
- Males: rho = -0.57, p = 0.042 \*
- Females: rho = -0.087, p = 0.506

#### sppb_velocidad_cat

```{r}
## Combined data
dat[, cor.test(hrv_pns_post, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_pns_post, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.107, p = 0.365
- Males: rho = 0.007, p = 0.981
- Females: rho = -0.135, p = 0.299

## Results 11: Greater SPPB with less RMSSD peri

Note: Using Spearman's rho as the constituents of the SPPB domains are ordinal (levels from 1 to 4) and RMSSD is continuous.

#### sppb_equil_cat

```{r}
## Combined data
dat[, cor.test(hrv_rmssd_peri, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_rmssd_peri, sppb_equil_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.085, p = 0.476
- Males: rho = 0.155, p = 0.614
- Females: rho = -0.179, p = 0.171

#### sppb_sts_cat

```{r}
## Combined data
dat[, cor.test(hrv_rmssd_peri, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_rmssd_peri, sppb_sts_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.327, p = 0.005 \*\*
- Males: rho = -0.419, p = 0.154
- Females: rho = -0.333, p = 0.009 \*\*

#### sppb_velocidad_cat

```{r}
## Combined data
dat[, cor.test(hrv_rmssd_peri, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      })]
## Separate analysis
dat[j = cor.test(hrv_rmssd_peri, sppb_velocidad_cat, method = "spearman", exact = F) |> 
      with({
        paste0("rho = ", round(estimate,3), ", p = ", round(p.value,3))
      }), 
    by = sd_sexo]
```

- Combined data: rho = -0.051, p = 0.669
- Males: rho = 0.283, p = 0.349
- Females: rho = -0.104, p = 0.426
